---
title: "Christmas 2024"
author: "Joey Stanley"
date: 2024-12-04
date-modified: last-modified
categories:
    - holidays
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")
```

Last year, I did [a detailed analysis](/posts/christmas) of Christmas hymns. This year, things have changed a little bit because of the introduction of some new hymns. So, in this post I'll give a brief look at what Christmas hymns are sung. It'll be similar in style to my ongoing post about [the second batch of new hymns](/posts/new_hymns_batch2), and I'll update it weekly until a negligible number of wards sing Christmas hymns. 

```{r}
december <- freqs |> 
    filter(date >= "2024-11-01",
           date <= today()) |> 
    print()

n_wards_here <- length(unique(december$ward))
n_weeks_here <- length(unique(december$date))
n_meetings_here <- length(unique(december$meeting_id))
```


```{r}
december |> 
    filter(wday(date, label = TRUE) != "Sun") |> 
    print()
```



```{r}
december |> 
    summarize(sang_new = sum(hymn_num %in% c(1018, 1201:1204)) > 0, .by = ward) |> 
    count(sang_new) |> 
    mutate(prop = n/sum(n))
```




```{r}
count_new_hymns <- function(.category_name, .hymns) {
    december |> 
        mutate(is_christmas = hymn_num %in% .hymns) |>
        summarize(n_christmas = sum(is_christmas), .by = c(date, ward)) |>
        mutate(n_christmas = if_else(n_christmas >= 2, 2, n_christmas)) |> 
        count(date, n_christmas) |>
        arrange(date) |>
        mutate(prop_christmas = n/sum(n), .by = date) |>
        mutate(n_christmas = as.factor(n_christmas)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_christmas, values_from = prop_christmas)
}
data_to_plot <- bind_rows(count_new_hymns("Old Christmas", c(201:214)), 
                          count_new_hymns("New Christmas", c(1018, 1202:1204)), 
                          count_new_hymns("Primary Christmas", c(2036:2054)),
                          count_new_hymns("Christmas",     c(201:214, 1018, 1202:1204, 2036:2054))) |> 
    pivot_longer(cols = -c(date, category), names_to = "n_christmas", values_to = "prop") |>
    mutate(group = paste(n_christmas, category)) |>
    arrange(date) |>
    # # don't need zeros from all three
    filter(!(n_christmas == 0 & category != "Christmas")) |>
    # don't need the 3s because they're so rare
    # filter(n_christmas <= 2) |>
    # mutate(n_christmas = if_else(n_christmas >= 2, 2, n_christmas)) |>
    filter(!is.na(prop)) |>
    print()
# Place labels a little above the max from most recent three weeks
# (Will need to raise to four as more dates come in.)
labels <- data_to_plot |> 
    slice_max(order_by = date, n = 3, by = n_christmas) |> 
    summarize(prop = max(prop), 
              date = max(date),
              .by = n_christmas) |> 
    mutate(label = case_when(n_christmas == 0 ~ "no Christmas hymns",
                             n_christmas == 1 ~ "1 Christmas hymn",
                             n_christmas >= 2 ~ "2 or more Christmas hymns"))
# Get date breaks for this post.
date_breaks <- tibble(weeks_since = 0:52,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= "2024-11-01")
```

First, what proportion of wards are singing Christmas hymns? The following plot shows the number of wards singing the number of Christmas hymsn per week. To make it easier to read, I've included the month of November, just so that there's a little bit of context. And yes, there were some wards that sang Christmas hymns in November, mostly wards outside the US since they don't celebrate Thanksgiving during that time, but also shout-out to a ward in Arizona that sang a Christmas hymn on November 17th! Anyway, the plot is a little confusing to look at, so let's see if I can explain it. 

<!--For now, I won't show a plot because it's only been one week, but I can summarize it in bullet points instead:

* 96.4% of wards sang at least one Christmas hymn.
* 12.5% of wards sang just one Christmas hymn.
* 83.5% of wards sang two or more Christmas hymns. (There was one ward that did three---yes, even on Fast Sunday---because they sang two closing hymns, both of which were Christmas. I've never seen that before!)
* 22.7% of wards sang one of the new hymns. The most common was <hymn>What Child Is This? (#1203)</hymn> by a fair margin, though see the figure below for some more detail. 
-->




```{r, include = TRUE, fig.height = 4, fig.width = 8}
ggplot(data_to_plot, aes(date, prop, group = group, color = group)) +
    # stat_smooth(formula = "y ~ x", method = "loess", color = "gray85", alpha = 0.25) +
    geom_path() +
    geom_point() +
    # geom_text(data = labels, aes(y = prop + 0.05, label = label), hjust = 1) +
    scale_x_date(breaks = date_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("0 Christmas",
                                  "1 Christmas", "1 Primary Christmas", "1 Old Christmas", "1 New Christmas",
                                  "2 Christmas", "2 Old Christmas", "2 New Christmas"),
                       labels = c("no Christmas hymns",
                                  "1 Christmas hymn (total)", "1 Primary Christmas", "1 old Christmas hymn", "1 new Christmas hymn",
                                 "2+ Christmas hymns (total)", "2+ old Christmas hymns","2+ new Christmas hymns"),
                       values = c("gray66", 
                                  "#a6cee3", "#fdbf6f", "#b2df8a", "#fb9a99",
                                  "#1f78b4", "#33a02c", "#e31a1c")) +
    theme_minimal() +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Wards singing Christmas hymns since November 1st",
         subtitle = paste("Based on data from", n_wards_here, "wards"),
         fill = "date")
    # theme(legend.position = "none")
```


First, we'll start with the gray line, which represents wards singing zero Christmas hymns. It's near the top during November and then quickly falls to the bottom. This means during November, the majorirty of wards I have data from sang zero Christmas hymns. But [as soon as December hit](/posts/Christmas), most wards sing at least one Christmas hymn. (But not all, as we'll see below!)

Let's move on to the blue lines, which represent the total number of Christmas hymns being sung. The darker line shows the number of wards that sing two or more Christmas hymns while the lighter line is for wards that sing exactly one. On December 1st, around 85% of wards sang two Christmas hymns, while just about 13% or so sang just one. Those numbers went up on December 8th and 15th.

The green, red, and orange lines give us a little insight into which hymns are sung. We'll start with the green lines. They're labeled "old" Christmas hymns, meaning they come from the 1985 hymnal. Most of the Christmas hymns come from this selection, which makes sense because there are more of them. Roughly two-thirds of wards sing two hymns from the existing hymnal with another 25% or so singing just one. With the red lines, we see that another 30% of wards sing one new hymn and a small number of wards sing two new hymns. Finally, I put the orange line in there because very occasionally, a ward will sing a Christmas song from the Children's Songbook. 


## How many Christmas hymns have wards sung so far?

Turning the question on its head then, we can now focus on individual wards and see how many Christmas hymns wards have sung so far.

```{r}
zero_christmas_wards <- december |> 
    mutate(is_christmas = hymn_num %in% c(201:214, 1018, 1202:1204, 2036:2054)) |> 
    summarize(n_christmas = sum(is_christmas), .by = c(meeting_id, date)) |> 
    filter(n_christmas == 0, 
           date >= "2024-12-01") |> 
    print()

december |> 
    filter(meeting_id %in% zero_christmas_wards$meeting_id) |> 
    arrange(date, ward, type)
```

First off, there were `r nrow(zero_christmas_wards)` sacrament meetings in December that I have data from that did not include any congregational Christmas hymns. One example was from a ward in Nevada. On December 8th, they sang <hymn>The Iron Rod (#274)</hymn>, <hymn>I Stand All Amazed (#193)</hymn>, and <hymn>I Believe in Christ (#134)</hymn>. I don't know why they didn't sing Christmas hymns well into December, but it just goes to show that it does happen very occasionally. 

The following plot shows an estimate of the cumulative number of Christmas hymns a ward sings over the course of the month of December. 

::: {.callout-note collapse="true"}
A little explanation of this plot might be necessary to fully understand it. When I collect my data, a lot of my data comes from the same wards each week, but a lot also comes from various other one-off contributions from people. So, the data for some wards is complete while for others it's not. That means I might have a complete dataset for some wards for the entire month and for other wards there might be just one or two contributions. 

So, the plot below shows that *of the wards I have just one week of data from*, around 81% of them sang two Christmas hymns. That doesn't mean that 81% of wards will necessarily sing two Christmas hymn on the first Sunday of the month. (Although, the plot above, I show that it's more like 85%.) It just means given one random week out of the month, the majority sing just two. A few wards will sing fewer than that and a few will sing more. 

See [here](/posts/new_hymns_batch1) for a similar plot, but applied to the first batch of new hymns.
:::


```{r, include = TRUE, fig.height = 7, fig.width = 4}
n_christmas_per_ward <- december |> 
    filter(date >= "2024-12-01") |> 
    summarize(n_weeks = length(unique(date)),
              n_christmas = sum(hymn_num %in% c(201:214, 1018, 1202:1204, 2036:2054)),
              .by = "ward") |>
    count(n_weeks, n_christmas) |> 
    mutate(prop_per_n = n/sum(n), .by = n_weeks) |> 
    mutate(prop_label = paste0(round(prop_per_n*100), "%"))

ggplot(n_christmas_per_ward, aes(n_weeks, n_christmas, color = prop_per_n)) + 
    geom_point(shape = 15, size = 9, position = position_nudge(x = -0.1)) + 
    geom_point(shape = 15, size = 9, position = position_nudge(x =  0.1)) +
    geom_text(aes(label = prop_label), color = "white") + 
    scale_x_continuous(breaks = 1:5,
                       expand = expansion(0, 0.5),
                       limits = c(0.5, 5.5)) + 
    scale_y_continuous(breaks = 0:18,
                       minor_breaks = NULL,
                       limits = c(0, 20),
                       expand = expansion(0, 0.5)) + 
    scale_color_distiller(palette = "Purples", direction = 1) + 
    labs(title = "Christmas hymns per ward in December",
         subtitle = "Percents refer to number of wards per column",
         x = "number of weeks I have data from, per ward",
         y = "number of Christmas hymns") + 
    theme_minimal() + 
    theme(legend.position = "none")
```

So, as expected, the more data I have from a ward, the more Christmas hymns I observe them singing. Currently, we're four weeks into December, and the most common pattern is for a ward to have sung between 8 and 10 Christmas hymns this year.




## What Christmas hymns have been most popular?

The following plot shows the estimated number of times each Christmas hymn is estimated to have been sung per ward this year. It's estimated that almost every ward sang <hymn>Joy to the World (#201)</hymn>. Meanwhile, less than about 20% of wards have sung <hymn>While Shepherds Watched Their Flocks (#211)</hymn>. 

```{r, include = TRUE, fig.height = 5, fig.width = 8}
cleveland_data <- freqs |> 
    mutate(color = case_when(hymn_num %in% c(201:214) ~ "old Christmas",
                             hymn_num %in% c(1018, 1202:1204) ~ "new Christmas",
                             TRUE ~ "other")) |> 
    filter(!hymn_num %in% c(91:94, 169:195, 1008),
           date >= "2024-12-01", 
           date <= last_sunday)
n_weeks_cleveland_data <- cleveland_data |> 
    distinct(date) |> 
    nrow()

cleveland_plot(cleveland_data, 
               return = "plot", 
               col_to_plot = avg_per_year,
               filter_n = 15,
               extra_cols = color,
               color = color,
               unique_meetings = get_n_distinct_meetings(cleveland_data),
               n_weeks = n_weeks_cleveland_data) + 
    scale_x_continuous(breaks = seq(0, 1.5, 0.1),
                       minor_breaks = seq(0, 1.5, 0.05),
                       # labels = scales::percent,
                       expand = expansion(0, c(0, 0.22))) +
    scale_color_manual(breaks = c("old Christmas", "new Christmas", "other"),
                       labels = c("Christmas hymns from the 1985 Hymnal", "Hymns from the New Hymnal", "other 1985 hymnal hymns"),
                       values = c("gray10", "#33a02c", "gray66")) + 
    labs(title = "Popularity of Christmas hymns in 2024",
         # subtitle = paste0("Based on ", n_meetings_here, " sacrament meetings from ", n_wards_here, " wards")) +
         x = "Estimated times sung this season",
         color = NULL
    ) + 
    theme_minimal() +
    theme(axis.text.y = element_blank(),
          legend.position = "bottom",
          legend.key.spacing.x = unit(1, "cm"))
```

Interspersed among these hymns are  new hymns. Of them, <hymn>What Child Is This? (#1203)</hymn> is the most popular; it is estimated to have been sung 0.6 times per ward, which means about 60% of wards sang it. The other three are closer to the bottom, more popular than a few of the ones in the 1985 hymnal, although <hymn>Come, Lord Jesus (#1018)</hymn> appears to have dethroned <hymn>While Shepherds Watched Their Glocks (#211)</hymn> for least popular Christmas hymn. I've added <hymn>Jesus, Once of Humble Birth (#196)</hymn> in there as well because it is far and away the most popular sacrament hymn around Christmas. This year, it was more common than half of the Christmas hymns.

The following table shows the same data plotted above but in table format so you can see the exact numbers.

```{r, include = TRUE}
cleveland_plot(cleveland_data, 
               return = "table", 
               col_to_plot = avg_per_year,
               filter_n = 4,
               extra_cols = color,
               color = color,
               unique_meetings = get_n_distinct_meetings(cleveland_data),
               n_weeks = n_weeks_cleveland_data) |> 
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    filter(hymn_num %in% c(196, 201:214, 1018, 1202:1204),) |> 
    mutate(percent_wards = scales::percent(round(n_weeks_here / every_x_weeks, 3))) |> 
    
    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           avg_per_year = paste0(round(avg_per_year * 100, 0), "%"),) |> 
    select(Hymn = label, `percent of wards` = avg_per_year) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards Christmas hymns have been sung in",
               subtitle = "Since December 1st")
```


## Comparison to previous years

This year, we introduced about four new Christmas hymns. Presumably, wards didn't simply add them to their existing programs. There are, after all, a finite number of hymns a ward can sing in December. So, by incorporating new Christmas hymns, some hymns had to get cut. So, which hymns got cut and were some hymns cut more than others? 

The following plot shows hymns in order of of how much more or less they were sung in 2024 compared to all the previous years I have data for. The way you read it is this: if the number is positive, it means the hymn was that much more common this year (e.g. "Away in a Manger was about 7% more common in 2024") but if it's negative, it means the hymn was that much less common this year (e.g. "While Shepherds Watched Their Flocks was a little over 40% less common in 2024").

```{r}
check_significance <- function(.hymn_num) {
    freqs |> 
        filter(month == 12) |> 
        mutate(is_2024 = year == 2024) |> 
        filter(!is.na(is_2024)) |> 
        mutate(is_211 = hymn_num == .hymn_num) |> 
        summarize(has_211 = sum(is_211) > 0, .by = c(meeting_id, is_2024)) |> 
    
        count(is_2024, has_211) |>
        pivot_wider(names_from = has_211, values_from = n) |>
            
            
        mutate(is_2024 = fct_recode(as.character(is_2024), 
                                    "in_2024" = "TRUE",
                                    "before_2024" = "FALSE")) |> 
        rename(meetings_sung = `TRUE`, meetings_not_sung = `FALSE`) |> 
        pivot_wider(names_from = is_2024, values_from = c(`meetings_sung`, `meetings_not_sung`)) |> 
        mutate(prop_meetings_2024 = meetings_sung_in_2024 / meetings_not_sung_in_2024,
               prop_meetings_before_2024 = meetings_sung_before_2024 / meetings_not_sung_before_2024,
               diff_in_2024 = prop_meetings_2024 / prop_meetings_before_2024)
}
check_significance(201)
diffs_2024 <- name_num_lookup |> 
    filter(hymn_num %in% c(196, 201:214)) |> 
    rowwise() |> 
    mutate(info = map(hymn_num, check_significance)) |> 
    unnest(info) |> 
    arrange(diff_in_2024) |> 
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           label = fct_inorder(label)) |> 
    print()
```

```{r, include = TRUE, fig.height = 4, fig.width = 8}
diffs_2024 |> 
    mutate(diff_in_2024 = diff_in_2024 - 1) |> 
    ggplot(aes(diff_in_2024, label, color = diff_in_2024)) + 
    geom_vline(xintercept = 0, color = "gray75", linetype = "dashed") + 
    geom_point(size = 5, color = "black") + 
    geom_point(size = 4) + 
    geom_text(aes(label = label), nudge_x = -0.03, color = "gray20", hjust = 1, size = 3) + 
    scale_x_continuous(breaks = seq(-0.5, 0.5, 0.1), 
                       labels = percent,
                       expand = expansion(0, c(0.25, 0.05))) + 
    scale_color_gradient2(high = "#117733", mid = "white", low = "red", midpoint = 0) + 
    labs(title = "How more/less common were Christmas hymns in 2024 compared to previous Decembers?",
         x = "percent more/less common") + 
    theme_minimal() + 
    theme(axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          legend.position = "none",
          axis.title.x = element_text(hjust = 0.74))
```

This is a pretty cool look at the characteristics of the 2024 season. A few of the more common hymns were a little bit more common this year. But, keep in mind that these are *congregational* hymns. It may not simply be that the most popular ones are getting even more popular. I think what's happening is that more wards choirs and other musical ensembles are singing/playing the new hymns, so in order to not hear the same Christmas hymn more than once per year, the music coordinator works around those. That means hymns that had been sung as choir pieces or as special musical numbers previously are now being shifted to the congregation. This is just speculation and since I haven't been collecting data on the special musical numbers themselves, I'll never know. But it's plausible. 

Most hymns were less common in 2024 than they were previously. They're the ones towards the right and bottom of the plot in red. I think these are the ones that are being left out in favor of the new hymns. Given the choice between <hymn>What Child Is This? (#1203)</hymn> and <hymn>While Shepherds Watched Their Flocks (#211)</hymn>, I'll bet most music coordinators went with the former. 

::: {.callout-caution collapse="true"}
### Statistical analysis for nerds :)

Is there anything special about that order? Maybe it's just the case that the new hymns were sung at the expense of the most popular hymns the least at the least popular hymns the most. In other words, perhaps there's a correlation between how popular a hymn is generally and how much *more* (or less) popular it was in 2024. Well, let's look at that correlation.

```{r, include = TRUE, fig.height = 4, fig.width = 8, warning = FALSE}
diffs_2024 |> 
    mutate(residual = lm(diff_in_2024 ~ prop_meetings_before_2024, data = diffs_2024)$residuals,
           diff_in_2024 = diff_in_2024 - 1) |> 
    ggplot(aes(prop_meetings_before_2024, diff_in_2024, color = residual)) +
    stat_smooth(method = "lm", formula = "y ~ x") + 
    geom_point(size = 4, color = "black") + 
    geom_point(size = 3) + 
    ggrepel::geom_text_repel(aes(label = hymn_name), color = "gray20", size = 3) + 
    scale_x_continuous(breaks = seq(0, 0.3, 0.04),
                       labels = percent) + 
    scale_y_continuous(breaks = seq(-0.5, 0.5, 0.25)) + 
    scale_color_gradient2(high = "#117733", mid = "white", low = "red", midpoint = 0) + 
    labs(title = "Difference in 2024 by general popularity of Christmas hymns",
         x = "percent sacrament meetings in December",
         y = "percent more/less common") + 
    theme_minimal() + 
    theme(legend.position = "none")
```

```{r}
cor_coef <- round(cor(diffs_2024$diff_in_2024, diffs_2024$prop_meetings_before_2024), 2)
mdl <- lm(diff_in_2024 ~ prop_meetings_before_2024, data = diffs_2024)
p_value <- round(summary(mdl)$coefficients[2,4], 3)
```

The plot shows that correlation is there (Ï = `r cor_coef`) and is statistically significant (*p* = `r p_value`). However, in this plot, I've colored the values based on how far off they are from where we'd expect them to be if there was a perfect correlation (for any statisticians reading this, it's their residuals). More popular hymns are towards the right and less popular are to the left (based on data prior to 2024). Meanwhile, how much more or less common a hymn was in 2024 is shown on the *y*-axis, with hymns sung far more than normal higher up and far less than normal towards the bottom. Points outside of the shaded area could be considered noteworthy, or rather, significantly more or less popular *given how popular that hymn has historically been*. 

Here we see that <hymn>Away in a Manger (#206)</hymn> is punching above its weight: it was sung far more in 2024 than you might expect solely based on how popular it has been in previous years. Same thing with <hymn>O Little Town of Bethlehem (#208)</hymn>. Meanwhile, <hymn>Far, Far Away on Judea's Plains (#212)</hymn> and <hymn>I Heard the Bells on Christmas Day (#214)</hymn> were sung quite a bit less this year even though it's about as popular as those two. 

So, this plot just shows that each hymn has its own story. It's not clear to me why these differences happened in 2024 the way they did. Perhaps it has something to do with the new hymns. But perhaps it's something else. (It could also be the case that if I ran the numbers for any year, I'd find big differences like these, which would mean 2024 is nothing special!) 

:::

## Conclusion

That's it for now! Hopefully this analysis of the 2024 Christmas season has been interesting. This is a special year because of the introduction of new Christmas hymns, so it's pretty cool to see how that has affected the trend for the past few years. 