---
title: "What are the altos' lowest notes?"
author: "Joey Stanley"
date: "03/03/2026"
date-modified: last-modified
categories:
  - soprano line
  - lowest notes
  - musicology
knitr:
  opts_chunk:
    out.width: "100%"
---

While my primary focus with this hymns stats blog has been to explore how the hymns are used in sacrament meetings, a secondary focus is to look at the music itself. I have already done a handful of posts on the music of the hymns, such as [the highest alto notes](/posts/highest_melody) and [the lowest soprano notes](/posts/lowest_soprano]. In this years-overdue post, we'll explore the lowest notes the alto line sings. Mark Mitchell has already done [fantastic work](https://markmitchell-music.com/music/documents/HymnStats.pdf) related to this. I hope to add to what he's reported already. 

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8,
                      freeze = TRUE)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")
alto <- notes %>% filter(voice == "alto")
```


Here's a plot that shows the distribution of lowest also notes across all the hymns in the 1985 hymnal.

```{r, include = TRUE, echo = FALSE, fig.height = 4}
lowest_notes_dist <- notes %>%
    filter(voice == "alto") %>%
    filter(midi == min(midi, na.rm = TRUE), .by = name_num) %>%
    distinct(name_num, note_label) %>%
    count(note_label)
ggplot(lowest_notes_dist, aes(note_label, n)) + 
    geom_col() + 
    scale_y_continuous(breaks = seq(0, 600, 20)) +
    labs(title = "Distributon of lowest notes in the alto line",
         x = "lowest note",
         y = "number of hymns") + 
    theme_minimal() + 
    theme(panel.grid.minor.x = element_blank())
```

As you can see there are three main categories of low notes: G3 and A♭3, A3 through B3, and C4 and above. Let's work through those three cateogires. 

## Lowest alto notes

### G3

Let's cut to the chase: the lowest note that an alto part has in the 1985 hymnal is a G3, which is the G below the treble clef. This occurs in eight different hymns for sure. Here they are in order of how common those hymns are sung as congregational hymns in sacrament meeting:

* Because I Have Been Given Much (#219)
* There Is Sunshine in My Soul Today (#227)
* In Fasting We Approach Thee (#139)
* Be Thou Humble (#130)
* Keep the Commandments (#303)
* Know This, That Every Soul Is Free (#240)
* God of Our Fathers, Known of Old (#80)
* Up, Awake, Ye Defenders of Zion (#248)


Several of these deserve some commentary. First, as mentioned I reported on the [lowest soprano notes](/posts/lowest_soprano), that G3 in <hymn>There Is Sunshine in My Soul Today (#227)</hymn> is sung with the lowest melody note that is not in unison with the alto. <hymn>Keep the Commandments (#303)</hymn> is a legitimate alto line, but it's only a duet, rather than a full SATB arrangement. The rest seem to be just simply lower-than-usual notes in otherwise normal arrangements.

There are two more in the [men's and women's arrangements](/posts/mens_womens_arrangements) in the back of the hymnal. One is in <hymn>Love at Home [Women] (#318)</hymn>, but I'm hesitant to treat that the same as the others because it's a SSA arrangement, so it's a little bit different than an SATB hymn. And the other is kinda not really an alto part in <hymn>The Priesthood of Our Lord [Men] (#320)</hymn>, which doesn't really count because it appears to be more of a unison hymn anyway, and that's if we should even consider a tenor harmony line an "alto" part. 

```{r}
alto %>%
    filter(midi == min(midi, na.rm = TRUE)) |> 
    pull(hymn_num) |> 
    single_hymn_lookup()
highest_lowest_note_stats("alto", "G3", "low")
```

Because some of these hymns are pretty common, singing this record-low note happens fairly often, a little less than once every three months on average. 

### A♭3

```{r}
alto %>%
    filter(midi == min(midi), .by = name_num) |> 
    filter(note_label == "Ab3") |> 
    pull(hymn_num) |> 
    single_hymn_lookup()
highest_lowest_note_stats("alto", "Ab3", "low")
```


The next lowest note, A♭3, is about as rare and is found in 10 hymns, including <hymn>I Stand All Amazed (#193)</hymn>, <hymn>As Now We Take the Sacrament (#169)</hymn>, and <hymn>A Poor Wayfaring Man of Grief (#29)</hymn>. Since the sacrament hymns that fall into this category are common and most of the non-sacrament hymns are not very common, it's primarily during the sacrament hymns that you see these low A♭s. 

## Typical lowest notes

So, we've established that G3 and A♭3 are unusually low. Looking at the distribution of the lowest notes across the hymnal in the plot above, let's move now on to the typical low end of alto lines, which is around an A3, B♭3, or B3. 

### A3

There are 79 hymns that have A3 as the lowest note for the altos. This includes sacrament hymns like <hymn>How Great the Wisdom and the Love (#195)</hymn>, other common hymns like <hymn>I Need Thee Every Hour (#98)</hymn>, and a surprising number of Christmas hymns, like <hymn>Hark! the Herald Angels Sing (#209)</hymn>. In fact, six of the fourteen Christmas hymns fall into this category, which makes singing A3 as the lowest note more common around Christmastime. The following chart shows this trend!

```{r}
highest_lowest_note_stats("alto", "A3", "low")
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "A3") |> 
    pull(hymn_num) |> 
    unique() |> 
    single_hymn_lookup()

alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "A3") |> 
    filter(hymn_num %in% 201:214) |> 
    count(name_num)
```


```{r, include = TRUE, fig.height = 4, fig.width = 8, echo = FALSE}
a3_hymns <- alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "A3") |> 
    pull(name_num) |> 
    unique()
freqs |> 
    filter(date <= ymd("2024-06-01")) |> 
    mutate(is_a3 = name_num %in% a3_hymns) |> 
    count(is_a3, week) |> 
    pivot_wider(names_from = is_a3, values_from = n) |> 
    mutate(prop = `TRUE` / (`FALSE` + `TRUE`)) |> 
    ggplot(aes(week, prop)) + 
    geom_path() + 
    geom_point() +
    scale_x_continuous(breaks = seq(0, 60, 6),
                       minor_breaks = seq(0, 60, 3),
                       expand = expansion(0, 1)) + 
    scale_y_continuous(labels = label_percent()) + 
    labs(title = "Percent of hymns sung where the lowest alto note is an A3",
         subtitle = "This seems to be especially common around Christmastime!",
         x = "week of the year",
         y = "percent of hymns sung") + 
    theme_minimal() 
```

The trend I think is purely coincidental. I don't think it says anything about culture, music, or anything of real significance. It's just interesting that a lot of Christmas hymns have A3 as the lowest alto note.

### B♭3

```{r}
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "Bb3") |> 
    filter(hymn_num %in% 169:196) |> 
    distinct(name_num)
highest_lowest_note_stats("alto", "Bb3", "low")
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "Bb3") |> 
    pull(hymn_num) |> 
    unique() |> 
    single_hymn_lookup()
```


Next are hymns where the lowest alto note is a B♭3. There are 64 hymns that fit into this category. Over half the time this type of hymn is sung is as a sacrament hymn since 9 of the 28 sacrament hymns have B♭3 as the lowest alto note, including <hymn>There Is a Green Hill Far Away (#194)</hymn> and <hymn>God Loved Us, So He Sent His Son (#187)</hymn>. This plot shows how often hymns where the lowest alto note is a B♭3 are sung:



```{r, include = TRUE, fig.height = 4, fig.width = 8}
Bb3_hymns <- alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "Bb3") |> 
    pull(name_num) |> 
    unique()
freqs |> 
    filter(date <= ymd("2024-06-01"),
           !is.na(type)) |> 
    mutate(is_bb3 = name_num %in% Bb3_hymns) |> 
    count(is_bb3, type) |> 
    pivot_wider(names_from = is_bb3, values_from = n) |> 
    mutate(prop = `TRUE` / (`FALSE` + `TRUE`)) |> 
    ggplot(aes(type, prop)) + 
    geom_col() + 
    scale_y_continuous(labels = label_percent()) + 
    labs(title = "Percent of hymns where the lowest alto note is a Bb3",
         subtitle = "This is as common as a sacrament hymn as the other slots combined!",
         x = NULL,
         y = "percent of hymns") +
    theme_minimal()
```

Again, I don't think this is a particularly meaningful insight and is probably nothing more than coincidence. But the size of that difference is pretty striking. Especially when compared to the distribution of A3 hymns, which is almost exactly even between Opening, Sacrament, and Closing hymns, with intermediate hymns being lower likely from the sheer fact that not all sacrament meetings have an intermediate hymn.

### B3

The most common lowest note for altos is a B3. This is found in 94 total hymns and makes up about a quarter of the hymns sung in sacrament meeting. These include some sacrament hymns like <hymn>Upon the Cross of Calvary (#184)</hymn> as well as numerous other common hymns like <hymn>Come, Come, Ye Saints (#30)</hymn>, <hymn>Choose the Right (#239)</hymn>, and <hymn>He Is Risen! (#199)</hymn>. There's nothing too interesting about the distribution of these hymns across the year or within the meeting.

```{r}
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "B3") |> 
    pull(hymn_num) |> 
    unique() |> 
    single_hymn_lookup()
highest_lowest_note_stats("alto", "B3", "low")
```

## Highest lowest notes

Now we enter the territory of the confusingly named "highest lowest notes". If we look at the lowest also notes across all hymns, some will be higher than others. This section looks at the ones that never go very low. If you're a soprano and have trouble singing below a middle C, for example, this might give you an idea of what hymns you could sing alto on.

### C4

```{r}
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "C4") |> 
    pull(hymn_num) |> 
    unique() |> 
    single_hymn_lookup()
highest_lowest_note_stats("alto", "C4", "low")
```

There are 25 hymns that have a C4 (middle C) as the lowest alto note. The most common one by a long shot in this group is the one sacrament hymn, <hymn>In Humility, Our Savior (#172)</hymn>. The next most common one is <hymn>I'll Go Where You Want Me to Go (#270)</hymn> which is sung only about every 16 months. The other hymns include several well-known hymns like <hymn>Put Your Shoulder to the Wheel (#252)</hymn>, <hymn>My Redeemer Lives (#135)</hymn>, and <hymn>Abide with Me; 'Tis Eventide (#165)</hymn>. On average, you're likely to see one of these hymns about every 5.6 weeks. 

### C♯4

```{r}
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "C#4") |> 
    pull(hymn_num) |> 
    unique() |> 
    single_hymn_lookup()
highest_lowest_note_stats("alto", "C#4", "low")
```

There are also 25 hymns that have C♯4 as the lowest alto note. None of these are sacrament hymns, but they do include more regular ones like <hymn>Lord, I Would Follow Thee (#220)</hymn>, <hymn>We Thank Thee, O God, for a Prophet (#19)</hymn>, and <hymn>Joy to the World (#201)</hymn>. In general, these are more common than the C4 hymns and on average you'll likely see one about every 4.2 weeks. 

### D4

```{r}
alto %>%
    filter(midi == min(midi), .by = name_num,
           note_label == "D4") |> 
    pull(hymn_num) |> 
    unique() |> 
    single_hymn_lookup()
highest_lowest_note_stats("alto", "D4", "low")
```

Finally, 11 hymns have the highest lowest alto note have alto lines that never go below a D4. <hymn>While of These Emblems We Partake (#174)</hymn> is the only sacrament hymn. Other include <hymn>Come, Ye Children of the Lord (#58)</hymn>, <hymn>It Came upon the Midnight Clear (#163)</hymn>, <hymn>Lord, Dismiss Us with Thy Blessing (#163)</hymn>, and <hymn>Abide with Me! (#166)</hymn>. I will show in later blog posts that some of these stand out in other ways as well: <hymn>Come, Ye Children of the Lord (#58)</hymn> generally has higher notes for everyone and <hymn>Lord, Dismiss Us with Thy Blessing (#163)</hymn> rather infamously has the most boring line ever (alto or otherwise). 

## Conclusion

In this post, I looked on the lowest notes for just the altos. I've shown that the lowest note that an alto would need to be able to sing in the hymnal is a G3. Some hymns though never go below a D4. Generally though, the lowest notes for altos is between A3 and Bb3.