---
title: "Overview"
author: "Joey Stanley"
date: 2023-10-23
date-modified: last-modified
---

```{r, include = FALSE}
rmarkdown::render("_scripts/analysis_functions.Rmd")
```

This page just has a summary of all the hymns and their frequency stats, including the new hymns. I will update this page regularly as I collect more data.

## Quick Facts

I currently have data from `r scales::comma(get_n_distinct_meetings())` sacrament meetings. If you were to attend that many, it would take roughly `r round(get_n_distinct_meetings()/48, 1)` years---so, more than a couple lifetimes' worth of data. It's also the equivalent of the entire 39-year lifetime of the 1985 hymnal in `r round(get_n_distinct_meetings()/48/39, 2)` wards. This data comes from `r scales::comma(length(unique(freqs$ward)))` wards, mostly in the United States. Across all these meetings, there are `r scales::comma(nrow(freqs))` hymns. See [here](/posts/frequency-data) for a detailed breakdown of how this data is distributed across time and space, as well as an explanation of where it call came from. 



## Frequency Table

::: {.panel-tabset}

### Before June 2024

This table includes data from before any new hymns were introduced. It is likely the most informative, publicly-available view of the overall popularity of hymns in the 1985 hymnal. 

```{r, include = TRUE, echo = FALSE}
freqs %>%
    filter(date <= ymd("2024-06-01")) |> 
    cleveland_plot(return = "table") %>%
    mutate(hymn_num = as.numeric(as.character(hymn_num))) %>%
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) %>%
    select(`#` = hymn_num,
           `Hymn` = hymn_name,
           `Times Sung` = n,
           `Avg per Year` = avg_per_year,
           `Every X Weeks` = every_x_weeks,
           `Every X Months` = every_x_months,
           `Every X Years`  = every_x_years) %>%
    mutate(across(c(`Avg per Year`:`Every X Years`), ~round(., 2))) %>%

    # Make ties the same rank
    rowid_to_column("Rank") %>%
    mutate(Rank = min(Rank), .by = `Times Sung`) %>%
# print()
    knitr::kable()
```


### Since June 2024

This table includes data from sacrament meetings since June 2024 when the new hymns were released. The numbers will be skewed quite a bit until about a full calendar year has passed. But it gives a decent idea of how popular the new hymns have been so far. 

```{r, include = TRUE, echo = FALSE}
freqs %>%
    filter(date >= ymd("2024-06-01")) |> 
    cleveland_plot(return = "table") %>%
    mutate(hymn_num = as.numeric(as.character(hymn_num))) %>%
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) %>%
    select(`#` = hymn_num,
           `Hymn` = hymn_name,
           `Times Sung` = n,
           `Avg per Year` = avg_per_year,
           `Every X Weeks` = every_x_weeks,
           `Every X Months` = every_x_months,
           `Every X Years`  = every_x_years) %>%
    mutate(across(c(`Avg per Year`:`Every X Years`), ~round(., 2))) %>%

    # Make ties the same rank
    rowid_to_column("Rank") %>%
    mutate(Rank = min(Rank), .by = `Times Sung`) %>%
# print()
    knitr::kable()
```

### All data

```{r, include = TRUE, echo = FALSE}
freqs %>%
    cleveland_plot(return = "table") %>%
    mutate(hymn_num = as.numeric(as.character(hymn_num))) %>%
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) %>%
    select(`#` = hymn_num,
           `Hymn` = hymn_name,
           `Times Sung` = n,
           `Avg per Year` = avg_per_year,
           `Every X Weeks` = every_x_weeks,
           `Every X Months` = every_x_months,
           `Every X Years`  = every_x_years) %>%
    mutate(across(c(`Avg per Year`:`Every X Years`), ~round(., 2))) %>%

    # Make ties the same rank
    rowid_to_column("Rank") %>%
    mutate(Rank = min(Rank), .by = `Times Sung`) %>%
# print()
    knitr::kable()
```

:::


## Visualization

Here is the same information, but as a plot.

```{r, echo = FALSE, fig.height = 40, fig.width = 8}
freqs %>%
    cleveland_plot(return = "plot", 
                   title = "Frequency of all hymns",
                   subtitle = NULL,
                   x_buffer = 0.7)
```

