---
title: "The Sixth Batch of New Hymns!"
author: "Joey Stanley"
date: "2026-03-05"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd(floor_date(today(), "week"))
# last_sunday <- batch6_release_date
library(gtExtras)
```




```{r}
general_conference_dates <- ymd("2024-10-06", "2025-04-06", "2025-10-04", "2025-10-05", "2026-04-05")
get_meetings_since_release <- function(.date) {
    freqs |> 
        filter(date >= ymd(.date),
               date <= ymd(last_sunday),
               !date %in% general_conference_dates)
}
meetings_since_release1 <- get_meetings_since_release(batch1_release_date)
meetings_since_release2 <- get_meetings_since_release(batch2_release_date)
meetings_since_release3 <- get_meetings_since_release(batch3_release_date)
meetings_since_release4 <- get_meetings_since_release(batch4_release_date)
meetings_since_release5 <- get_meetings_since_release(batch5_release_date)
meetings_since_release6 <- get_meetings_since_release(batch6_release_date)

get_n_meetings_since_release <- function(.df) {nrow(distinct(.df, meeting_id))}
n_meetings_since_release1 <- get_n_meetings_since_release(meetings_since_release1)
n_meetings_since_release2 <- get_n_meetings_since_release(meetings_since_release2)
n_meetings_since_release3 <- get_n_meetings_since_release(meetings_since_release3)
n_meetings_since_release4 <- get_n_meetings_since_release(meetings_since_release4)
n_meetings_since_release5 <- get_n_meetings_since_release(meetings_since_release5)
n_meetings_since_release6 <- get_n_meetings_since_release(meetings_since_release6)

get_n_wards_since_release <- function(.df) {
    .df |> 
        filter(date <= ymd(last_sunday)) |> 
        pull(ward) |> 
        unique() |> 
        length()
}
n_wards_since_release1 <- get_n_wards_since_release(meetings_since_release1)
n_wards_since_release2 <- get_n_wards_since_release(meetings_since_release2)
n_wards_since_release3 <- get_n_wards_since_release(meetings_since_release3)
n_wards_since_release4 <- get_n_wards_since_release(meetings_since_release4)
n_wards_since_release5 <- get_n_wards_since_release(meetings_since_release5)
n_wards_since_release6 <- get_n_wards_since_release(meetings_since_release6)

get_n_weeks_since_release <- function(.df) {nrow(count(.df, date))}
n_weeks_since_release1 <- get_n_weeks_since_release(meetings_since_release1)
n_weeks_since_release2 <- get_n_weeks_since_release(meetings_since_release2)
n_weeks_since_release3 <- get_n_weeks_since_release(meetings_since_release3)
n_weeks_since_release4 <- get_n_weeks_since_release(meetings_since_release4)
n_weeks_since_release5 <- get_n_weeks_since_release(meetings_since_release5)
n_weeks_since_release6 <- get_n_weeks_since_release(meetings_since_release6)
```


On February 26, 2026, the church [released](https://newsroom.churchofjesuschrist.org/article/old-testament–themed-hymns-join-hymns-for-home-and-church) the sixth batch of new hymns. In previous posts, I have covered in detail the previous batches of hymns and how they were incorporated into sacrament meetings. This page covers period since when the sixth batch came out until now.

Currently I have data from `r scales::comma(n_meetings_since_release6)` sacrament meetings from `r scales::comma(n_wards_since_release6)` wards since February 26, 2026. Here's how much data I have for each week so far, just so you have an idea of what I'm working with.

```{r, include = TRUE}
meetings_since_release6 |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since Febrary 26, 2026")
```


```{r}
# show all sixth batch data
meetings_since_release6 |> 
    filter(hymn_num %in% batch6_hymn_nums)
```


## How many wards sang new hymns each week?

The following plot show what percentage of wards sang from each of the new batches of hymns, per week, since June 2024 when the new hymns started getting rolled out. I've included data from all batches just so you can compare them to each other.

```{r, fig.height = 4, fig.width = 8}
count_new_hymns <- function(.category_name, .hymns) {
    meetings_since_release1 |> 
        mutate(is_new = hymn_num %in% .hymns) |>
        summarize(n_new = sum(is_new), .by = c(date, ward)) |>
        count(date, n_new) |>
        arrange(date) |>
        mutate(prop_new = n/sum(n), .by = date) |>
        mutate(n_new = as.factor(n_new)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_new, values_from = prop_new)
}
data_to_plot <- meetings_since_release1 |> 
    summarize(batch1 = sum(hymn_num %in% batch1_hymn_nums) > 0,
              batch2 = sum(hymn_num %in% batch2_hymn_nums) > 0,
              batch3 = sum(hymn_num %in% batch3_hymn_nums) > 0,
              batch4 = sum(hymn_num %in% batch4_hymn_nums) > 0,
              batch5 = sum(hymn_num %in% batch5_hymn_nums) > 0,
              batch6 = sum(hymn_num %in% batch6_hymn_nums) > 0,
              .by = c(meeting_id, date)) |>
    mutate(none = !batch1 & !batch2 & !batch3 & !batch4 & !batch5 & !batch6) |>
    pivot_longer(cols = c(batch1, batch2, batch3, batch4, batch5, batch6, none), names_to = "batch", values_to = "present") |> 
    count(date, batch, present) |> 
    pivot_wider(names_from = present, values_from = n, values_fill = 0) |> 
    mutate(prop_sung = `TRUE`/(`TRUE` + `FALSE`)) |> 
    
    # don't need accidental data from before the release date
    filter(!(batch == "batch6" & date < ymd(batch6_release_date)),
           !(batch == "batch5" & date < ymd(batch5_release_date)),
           !(batch == "batch4" & date < ymd(batch4_release_date)),
           !(batch == "batch3" & date < ymd(batch3_release_date)),
           !(batch == "batch2" & date < ymd(batch2_release_date)),
           !(batch == "batch1" & date < ymd(batch1_release_date))) |>
    print()
```



```{r, fig.height = 4, fig.width = 8}
weeks_since_initial_release <- ceiling(as.numeric(last_sunday - ymd(batch1_release_date)) / 7)

# Get date breaks for this post.
date_minor_breaks <- tibble(weeks_since = 0:weeks_since_initial_release,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= batch1_release_date)
date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 6 == 0)
batch6_date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 2 == 0)
```


```{r}
labels <- tibble(
    batch = c("none", "batch1", "batch2", "batch3", "batch4", "batch5"),
    prop  = c(0.73, 0.33, 0.08, 0.04, 0.04, 0.01),
    date  = ymd(c("2025-03-25", "2024-08-11", "2025-01-25", "2025-03-15", "2025-06-15", "2025-10-20")),
    label = c("no new hymns", "batch 1", "batch 2", "batch 3", "batch4", "batch5"))

# Annotations
christmas2024_data_to_highlight <- data_to_plot |> 
    filter(!batch %in% c("none"),
           date >= "2024-12-01",
           date <= "2025-01-01") |> 
    mutate(label = "Christmas")
christmas2025_data_to_highlight <- data_to_plot |> 
    filter(!batch %in% c("none"),
           date >= "2025-12-01",
           date <= "2026-01-01") |> 
    mutate(label = "Christmas")
easter_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2025-04-13",
           date <= "2025-04-27") |> 
    mutate(label = "Easter")
```


```{r}
p <- ggplot(data_to_plot, aes(date, prop_sung, group = batch, color = batch)) +
    
    # Christmas annotations
    ggforce::geom_mark_ellipse(data = christmas2024_data_to_highlight, group = 1, color = "gray50") + 
    annotate(geom = "text", x = ymd("2024-11-03"), y = 0.35, label = "Christmas", hjust = 1, color = "gray50") + 
    annotate(geom = "curve", color = "gray50",
             x = ymd("2024-11-05"), xend = ymd("2024-11-25"), y = 0.35, yend = 0.30,
             curvature = -0.25, arrow = arrow(length = unit(2, "mm"))) + 
    
    ggforce::geom_mark_ellipse(data = christmas2025_data_to_highlight, group = 1, color = "gray50") + 
    annotate(geom = "text", x = ymd("2025-11-03"), y = 0.35, label = "Christmas", hjust = 1, color = "gray50") + 
    annotate(geom = "curve", color = "gray50",
             x = ymd("2025-11-05"), xend = ymd("2025-11-25"), y = 0.35, yend = 0.30,
             curvature = -0.25, arrow = arrow(length = unit(2, "mm"))) + 
    
    # Easter
    # ggforce::geom_mark_ellipse(data = easter_data_to_highlight, group = 1, color = "gray50") +
    # annotate(geom = "segment", color = "gray50",
    #          x = ymd("2025-04-20"), xend = ymd("2025-04-20"), y = 0.35, yend = 0.275, 
    #          arrow = arrow(length = unit(2, "mm"))) +
    # annotate(geom = "text", x = ymd("2025-04-20"), y = 0.37, label = "Easter", hjust = 0.5, color = "gray50") +
    # 
    # # Mother's Day
    # annotate(geom = "segment", 
    #          x = ymd("2025-05-11"), xend = ymd("2025-05-11"), y = 0.28, yend = 0.18, 
    #          arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    # annotate(geom = "text", x = ymd("2025-05-11"), y = 0.29, label = "Mother's\nDay", 
    #          vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +
    # 
    # # Father's Day
    # annotate(geom = "segment", 
    #          x = ymd("2025-06-15"), xend = ymd("2025-06-15"), y = 0.28, yend = 0.18, 
    #          arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    # annotate(geom = "text", x = ymd("2025-06-15"), y = 0.29, label = "Father's\nDay", 
    #          vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +
    
    
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(data = labels, aes(y = prop, label = label), hjust = 0.5, vjust = 0.5, size = 6) +
    
    
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4", "batch5", "batch6"),
                       # Polychrome::kelly.colors
                       values = c("#848482", "#be0032",  "#0067a5",  "#008856", "#875692", "#f38400", "#8db600")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since June 2, 2024",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```


::: {.panel-tabset}

### The last six months

```{r, include = TRUE, fig.height = 5.5, fig.width = 14, warning = FALSE}
#| column: screen
six_months_ago <- last_sunday - round(365/2)

p + 
    labs(title = paste0("Percentage of wards singing new hymns since ", sf(six_months_ago), ", ", year(six_months_ago))) + 
    coord_cartesian(xlim = c(ymd(batch5_release_date), ymd(last_sunday)))
```
### Since the new hymns came out

```{r, include = TRUE, fig.height = 5.5, fig.width = 14}
#| column: screen
p
```
:::


The sixth batch (highlighted in yellowish-green) took a few months to get incorporated into sacrament meetings as a congregational hymn. For the first two months, it was extremely rare to see it sung. In late April and into May, it picked up a little bit. I suspect that whoever plans music in wards had six or so weeks planned out already when the batch came out, so it was sung more when those preplanned ones had already passed. However, it's worth noting that every batch since the third batch took a few weeks to get warmed up, and each batch seems to be less popular than the one previous. But the sixth batch took an especially long time to get going.

Since then though, it has done somewhat well. It is about as popular as the fourth and fifth batches. It's still not as popular as the second and third and certainly not the first batch. One thing we see though it a small spike on Fast Sundays, which is from <hymn>`r hymn(1062)`</hymn> being sung then. In fact, if you look at the gray line, which represents the percent of wards that sang zero new hymns each week, for the past two years there is a spike *up* on Fast Sunday as people continue to sing the typical [Fast Sunday](/posts/fast_sunday) hymns. That might be going away a little bit now that we clearly have a new hymn to throw into the mix.


It's worth pointing one more thing about that gray line. Since about October 2025, it's is mostly staying under the 50% mark. That means that on a given Sunday, there are more wards that sing a new hymn than there are that don't. The steady decline from June 2024 to now I think is pretty remarkable. The lowest point was on May 17, 2026 when 56.4% of wards sang at least one new hymn. It takes a few months for a batch to ramp up to its full usage, and none have really changed in popularity all that much over time. That means that we're steadily incorporating more new hymns as the batches come out. If we're singing a Batch 1 hymn every three or four weeks, we'll sprinkle in a second and third batch one every five weeks. And when a new batch comes out, it gets slid into remaining slots. The result is fewer weeks without any new hymns. It also seems like no batches are "eating into" previous ones. I think we really are just adding to the new hymns we're already singing.


```{r}
# show all sixth batch data
meetings_since_release6 |> 
    filter(hymn_num %in% batch6_hymn_nums)

# by hymn
meetings_since_release6 |> 
    filter(hymn_num %in% batch6_hymn_nums) |> 
    count(hymn_name, hymn_num, sort = TRUE) |> 
    mutate(prop = n/sum(n))
# by hymn and by date
meetings_since_release6 |> 
    filter(hymn_num %in% batch6_hymn_nums) |> 
    count(hymn_name, hymn_num, date, sort = TRUE) |> 
    mutate(prop = n/sum(n), .by = date)

# by date
data_to_plot |> 
    filter(batch == "batch6")
# sorted by best weeks
data_to_plot |> 
    filter(batch == "batch6") |> 
    arrange(-prop_sung)
# fourth batch for comparison
data_to_plot |> 
    filter(batch == "batch5") |> 
    arrange(-prop_sung)

# gray line
data_to_plot |> 
    filter(batch == "none") |> 
    arrange(prop_sung) |> 
    mutate(1-prop_sung)
```

Here's another version of the same plot but instead of the actual dates, it's in terms of how many weeks since the batches came out. This version of the plot is useful to spot any parallel trends across the batches, especially in the first few weeks of their release.

```{r, include = TRUE, fig.height = 4.5, fig.width = 14}
#| column: page
data_to_plot |> 
    mutate(weeks_since_release = as.numeric(date - min(date))/7, .by = batch) |> 
    filter(!batch %in% "none") |> 
    ggplot(aes(weeks_since_release, prop_sung, group = batch, color = batch)) +
    geom_path(linewidth = 1) +
    geom_point(size = 2) +
    scale_x_continuous(breaks = seq(2, 150, 4),
                       expand = expansion(0, c(0.2, 2))) + 
    scale_y_continuous(expand = expansion(0, c(0.01, 0.02)),
                       breaks = seq(0, 1, 0.1),
                       minor_breaks = seq(0, 1, 0.02),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4", "batch5", "batch6"),
                       # Polychrome::kelly.colors
                       values = c("#848482", "#be0032",  "#0067a5",  "#008856", "#875692", "#f38400", "#8db600")) +
    labs(x = "weeks since release",
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since their release date",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "bottom")
```


Here it's easier to see that the first two months of the sixth batch have been the least popular out of any batch so far. It has gotten about as popular as the other batches since then, but it still lags just a little bit.


## What hymns are most popular?

The following table shows the hymns from the sixth batch in order of how popular they are.  

```{r, include = TRUE}
meetings_since_release6 |> 
    
    # I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |> 
    filter(nth_time == 1) |> 
    
    mutate(hymn_num = convert_to_2000(hymn_num)) |> 
    
    cleveland(return = "table", 
              stat_to_plot = sum_prop_wards,
              hymns_to_plot = batch6_hymn_nums) |> 
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    filter(hymn_num %in% batch6_hymn_nums) |>
    mutate(sum_prop_wards = scales::percent(round(sum_prop_wards, 3))) |> 
    
    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           sum_prop_wards = if_else(is.na(sum_prop_wards), "0%", sum_prop_wards)) |> 
    select(Hymn = label, `percent of wards` = sum_prop_wards) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Since February 26, 2026")
```

The two most popular hymns in this batch are <hymn>`hymn(1062)`</hymn> and <hymn>`hymn(1061)`</hymn>. The first one is explained because it's a Fast Sunday hymn so there's a good reason to sing it every month. The second was popular around [Father's Day](/posts/fathers_day) and [Mother's Day](/posts/mothers_day). While not apparent in this table, <hymn>`r hymn(1210)`</hymn> was also somewhat popular around [Easter](/posts/easter). It was brand new then, so it didn't get sung much, but I suspect next year we'll see it sung more.  It's pretty cool to see new hymns being introduced in those categories since the number of hymns in each one of those is pretty small.

## What hymns are most popular each week?

This plot shows how popular each hymn is each week. Take these numbers with a grain of salt, particularly the left sides of each panel, because there's not a lot of data to work with. 

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 13}
#| column: page
meetings_since_release6 |>
     filter(hymn_num %in% batch6_hymn_nums) |>
     count(date, hymn_num, name_num) |>

     # pivot and stuff to get zeros
     pivot_wider(names_from = date, values_from = n, values_fill = 0) |>
     pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
     mutate(date = ymd(date)) |>

     # put them in order
     arrange(hymn_num) |>
     mutate(name_num = fct_inorder(name_num)) |>

     # Put the dates in order
     arrange(date) |>
     mutate(date = date) |>

     # proportions per week
     mutate(prop = n/sum(n), .by = date) |>

     ggplot(aes(date, prop, group = name_num)) +
     geom_path() +
     geom_point() +
     scale_x_date(date_breaks = "3 weeks",
                  date_minor_breaks = "1 week",
                  date_labels = "%b %d",
                  expand = expansion(0.1, 0)) +
     scale_y_continuous(labels = scales::percent) +
     facet_wrap(~name_num, ncol = 4) +
     labs(title = "Proportion of wards singing new hymns by week",
          subtitle = "Of the sacrament meetings with new hymns",
          x = NULL,
          y = "proportion of wards") +
     theme_bw()
```


The clearest trend is the rise and fall of <hymn>`r hymn(1210)`</hymn> around Easter. The other one is the regular spike in <hymn>`r hymn(1062)`</hymn> each month. We can also see <hymn>`r hymn(1061)`</hymn> spiking around Mother's Day and Father's Day. The other hymns don't show too clear of a pattern. 

## When during meetings are these hymns sung?

This table shows when during sacrament meetings these hymns have been sung.

```{r, include = TRUE}
meetings_since_release6 |>
    mutate(hymn_num = convert_to_2000(hymn_num)) |> 
    filter(hymn_num %in% batch6_hymn_nums,
           !is.na(type)) |>
    count(name_num, hymn_num, type) |>
    mutate(n = round(n/sum(n), 2)*100,
           .by = name_num) |>
    pivot_wider(names_from = type, values_from = n, values_fill = 0) |>
    arrange(hymn_num) |>
    select(-hymn_num) |>
    select(name_num,
           Opening, Sacrament, Intermediate, Closing) |>
    rename(`New Hymn` = name_num) |>
    gt() |>
    gt_plt_bar_pct(column = c(Opening), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Sacrament), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Intermediate), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Closing), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Since February 26, 2026") |>
    tab_source_note("Note: Each row adds up to 100%.")
```


Like we've seen with other batches and pretty much any other hymn, each hymn has its own story. I don't know if there's much to be said about this data. 





## Conclusion

It took a few weeks to get started, but the sixth batch is being incorporated into sacrament meetings. Since we have some topical hymns, it'll take a few years for the true popularity of them to be evened out across the months and years. In general though, it does seem to be the least popular batch so far.

This is just a continuation of the trend seen since the second batch came out---newer batches are less popular. I've said this before in previous analyses, but it might be because the newness is losing its excitement. It might also be that the hymnal committee purposely chose hymns they anticipated being popular at the beginning. I'm thinking ahead to when the full hymnal gets released because presumably we'll get a whole bunch of new ones then. I'm curious if those new ones will ever become as popular as the ones released in these batches. 