---
title: "The Third Batch of New Hymns!"
author: "Joey Stanley"
date: "2025-02-16"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd(floor_date(today(), "week"))
library(gtExtras)
```



```{r}
meetings_since_release1 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-06-02"),
           !date %in% ymd("2024-10-06",
                        "2025-04-06"))
meetings_since_release2 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-09-15"),
           !date %in% ymd("2024-10-06",
                        "2025-04-06"))
meetings_since_release3 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2025-02-16"),
           !date %in% ymd("2025-04-06"))

n_meetings_since_release1 <- meetings_since_release1 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release2 <- meetings_since_release2 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release3 <- meetings_since_release3 |> 
    distinct(meeting_id) |> 
    nrow()

n_wards_since_release1 <- meetings_since_release1 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release2 <- meetings_since_release2 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release3 <- meetings_since_release3 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()

n_weeks_since_release1 <- meetings_since_release1 |> 
    count(date) |> 
    nrow()
n_weeks_since_release2 <- meetings_since_release2 |> 
    count(date) |> 
    nrow()
n_weeks_since_release3 <- meetings_since_release3 |> 
    count(date) |> 
    nrow()
```


On February 13th, 2025, the church [released](https://newsroom.churchofjesuschrist.org/article/fifteen-new-hymns-available-to-use-in-church-and-at-home) the third batch of new hymns. In previous posts, I have covered in detail the [first](/posts/new_hymns_batch1) and [second](/posts/new_hymns_batch2) batches of hymns and how they were incorporated into sacrament meetings. This page covers period since when the third batch came out. I'll update this page weekly as I collect more data. Currently I have data from `r scales::comma(n_meetings_since_release3)` sacrament meetings from `r scales::comma(n_wards_since_release3)` wards since February 16th.

But first, here's how much data I have for each week so far, just so you have an idea of what I'm working with.

```{r, include = TRUE}
meetings_since_release3 |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since February 16th")
```



## How many wards sang new hymns each week?

The following plot show what percentage of wards sang from each of the new batches of hymns, per week, since June 2024 when the new hymns started getting rolled out. I've included data from all three batches just so you can compare them to each other.

```{r, fig.height = 4, fig.width = 8}
count_new_hymns <- function(.category_name, .hymns) {
    meetings_since_release1 |> 
        mutate(is_new = hymn_num %in% .hymns) |>
        summarize(n_new = sum(is_new), .by = c(date, ward)) |>
        count(date, n_new) |>
        arrange(date) |>
        mutate(prop_new = n/sum(n), .by = date) |>
        mutate(n_new = as.factor(n_new)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_new, values_from = prop_new)
}
data_to_plot <- meetings_since_release1 |> 
    summarize(batch1 = sum(hymn_num %in% c(1001:1009, 1201:1204)) > 0,
              batch2 = sum(hymn_num %in% c(1010:1018)) > 0,
              batch3 = sum(hymn_num %in% c(1019:1031, 1205:1206)) > 0,
              .by = c(meeting_id, date)) |>
    mutate(none = !batch1 & !batch2 & !batch3) |>
    pivot_longer(cols = c(batch1, batch2, batch3, none), names_to = "batch", values_to = "present") |> 
    count(date, batch, present) |> 
    pivot_wider(names_from = present, values_from = n, values_fill = 0) |> 
    mutate(prop_sung = `TRUE`/(`TRUE` + `FALSE`)) |> 
    
    filter(!(batch == "batch3" & date < ymd("2025-02-13"))) |>
    # don't need batch 2 before Sep 15
    filter(!(batch == "batch2" & date < ymd("2024-09-15"))) |>
    # don't need batch 1 before Sep 15 either
    filter(!(batch == "batch1" & date < ymd("2024-06-02"))) |>
    print()


# Place labels a little above the max from most recent three weeks
# (Will need to raise to four as more dates come in.)
# labels <- data_to_plot |> 
#     slice_max(order_by = date, n = 3, by = batch) |> 
#     summarize(prop = max(prop_sung), 
#               date = max(date),
#               .by = batch) |> 
#     mutate(label = case_when(batch == "none" ~ "no new hymns",
#                              batch == "batch1" ~ "batch 1",
#                              batch == "batch2" ~ "batch 2",
#                              batch == "batch3" ~ "batch 3"))

```



```{r, fig.height = 4, fig.width = 8}
# Get date breaks for this post.
date_minor_breaks <- tibble(weeks_since = 0:52,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= "2024-06-02")
date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 3 == 0)
batch3_date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 2 == 0)
```


```{r}
labels <- tibble(
    batch = c("none", "batch1", "batch2", "batch3"),
    prop  = c(0.73, 0.26, 0.08, 0.04),
    date  = ymd(c("2025-04-01", "2025-02-14", "2025-01-25", "2025-03-15")),
    label = c("no new hymns", "batch 1", "batch 2", "batch 3"))

# Annotations
christmas_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2024-12-01",
           date <= "2025-01-01") |> 
    mutate(label = "Christmas")
easter_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2025-04-13",
           date <= "2025-04-27") |> 
    mutate(label = "Easter")
```


```{r, include = TRUE, fig.height = 5.5, fig.width = 14}
#| column: page
ggplot(data_to_plot, aes(date, prop_sung, group = batch, color = batch)) +
    
    # Christmas annotation
    ggforce::geom_mark_ellipse(data = christmas_data_to_highlight, group = 1, color = "gray50") + 
    annotate(geom = "text", x = ymd("2024-11-03"), y = 0.35, label = "Christmas", hjust = 1, color = "gray50") + 
    annotate(geom = "curve", color = "gray50",
             x = ymd("2024-11-05"), xend = ymd("2024-11-25"), y = 0.35, yend = 0.30,
             curvature = -0.25, arrow = arrow(length = unit(2, "mm"))) + 
    
    # Easter
    ggforce::geom_mark_ellipse(data = easter_data_to_highlight, group = 1, color = "gray50") +
    annotate(geom = "segment", color = "gray50",
             x = ymd("2025-04-20"), xend = ymd("2025-04-20"), y = 0.35, yend = 0.275, 
             arrow = arrow(length = unit(2, "mm"))) +
    annotate(geom = "text", x = ymd("2025-04-20"), y = 0.37, label = "Easter", hjust = 0.5, color = "gray50") +
    
    # Mother's Day
    annotate(geom = "segment", 
             x = ymd("2025-05-11"), xend = ymd("2025-05-11"), y = 0.28, yend = 0.18, 
             arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    annotate(geom = "text", x = ymd("2025-05-11"), y = 0.29, label = "Mother's\nDay", 
             vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +
    
    
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(data = labels, aes(y = prop, label = label), hjust = 0.5, vjust = 0.5, size = 6) +
    
    
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3"),
                       values = c("gray66", "#1b9e77",  "#d95f02",  "#7570b3")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since June 2",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards"),
         fill = "date") + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

```{r}
as_percent_joey <- function(x){
    paste0(round(x * 100, 1), "%")
}
n_meetings_on_June2 <- meetings_since_release1 |> 
    filter(date == "2024-06-02") |> 
    count(meeting_id) |> 
    nrow()
prop_new_batch1_week1 <- data_to_plot |> 
    filter(date == "2024-06-02",
           batch == "batch1") |>
    pull(prop_sung) |> 
    as_percent_joey()
```

In June, the first batch was readily incorporated into congregational singing. That first week, about `r prop_new_batch1_week1` of the `r n_meetings_on_June2` wards I have data from sang one of the new hymns. It stayed pretty consistently about that high for the next three months. During Christmastime, it surged in popularity again, mostly because there were some Christmas hymns introduced in that batch.

```{r}
prop_new_batch2_week1 <- data_to_plot |> 
    filter(date == "2024-09-15",
           batch == "batch2") |>
    pull(prop_sung) |> 
    as_percent_joey()
prop_new_batch2_week2 <- data_to_plot |> 
    filter(date == "2024-09-22",
           batch == "batch2") |>
    pull(prop_sung) |> 
    as_percent_joey()
prop_new_batch2_Oct20 <- data_to_plot |> 
    filter(date == "2024-10-20",
           batch == "batch2") |>
    pull(prop_sung) |> 
    as_percent_joey()
```

When the second batch came out, only about `r prop_new_batch2_week1` of wards sang at least one of those hymns right away. But, that jumped up to `r prop_new_batch2_week2` and eventually peaked five weeks in at `r prop_new_batch2_Oct20` on October 20. It took a nosedive during Christmastime since there weren't any Christmas hymns in that batch. After that, it partially recovered in January and early February, hovering around 13% of wards singing from it each week. Meanwhile, the first batch of hymns has almost always been more popular. In October 2024, when the second batch was most popular, the two were about even. Since Christmas, the first batch has consistently been more popular than the second batch.

```{r}
n_meetings_on_Feb16 <- meetings_since_release1 |> 
    filter(date == "2025-02-16") |> 
    count(meeting_id) |> 
    nrow()
prop_new_batch3_week1 <- data_to_plot |> 
    filter(date == "2025-02-16",
           batch == "batch3") |>
    pull(prop_sung) |> 
    as_percent_joey()
prop_new_batch3_week2 <- data_to_plot |> 
    filter(date == "2025-02-23",
           batch == "batch3") |>
    pull(prop_sung) |> 
    as_percent_joey()
prop_new_batch3_week3 <- data_to_plot |> 
    filter(date == "2025-03-02",
           batch == "batch3") |>
    pull(prop_sung) |> 
    as_percent_joey()
prop_new_batch3_week7 <- data_to_plot |> 
    filter(date == "2025-04-13",
           batch == "batch3") |>
    pull(prop_sung) |> 
    as_percent_joey()
```

Now we get to the third batch. If the second batch's reception was lukewarm, the third batch was downright cold. On the first week of the third batch, I have data from `r n_meetings_on_Feb16` wards and just `r prop_new_batch3_week1` of them sang one of the brand new songs. The second week was better, with `r prop_new_batch3_week2`, but still far below where the other batches were during their first few weeks. It's not like people are no longer singing new hymns; in fact, the first and second batches are going as strong as ever. In the past few weeks, the third batch has finally surpassed the second batch in weekly popularity, partly due to <hymn>Let Easter Anthems Ring (#1205)</hymn> and <hymn>Were You There? (#1206)</hymn> around Easter, and then on Mother's Day, it passed the first batch. Although, I have markedly less data from Mother's Day, so I'll have to get more to see if that holds true.

There may be a few reasons why fewer wards are singing from this third batch. At first, perhaps music coordinators had the hymns planned out for the month or quarter already and didn't want to change plans, particularly for the sake of the organists. Maybe the novelty of new hymns has worn off a little bit. I *would* say maybe because there weren't any "bangers" like <hymn>Come, Thou Fount of Every Blessing (#1001)</hymn> and <hymn>Amazing Grace (#1010)</hymn>, but I'd put <hymn>This Little Light of Mine (#1028)</hymn> and <hymn>Faith in Every Footstep (#1022)</hymn> up there. Regardless, the point of the matter is that the third batch has not been as popular so far.

```{r}
data_to_plot |> 
    filter(batch == "none") |> 
    arrange(prop_sung) |> 
    print()
```




## What hymns are most popular?

The following table shows the hymns from the third batch in order of how popular they are. On the right, I have an approximation of how many wards have sung that hymn since its release. 

```{r, include = TRUE}
meetings_since_release1 |> 
    
    # Because people are repeating 1001 now, I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |> 
    filter(nth_time == 1) |> 
    
    cleveland(return = "table", 
              stat_to_plot = sum_prop_wards,
              hymns_to_plot = c(1019:1031, 1205:1206)) |> 
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    filter(hymn_num %in% c(1019:1031, 1205:1206)) |>
    mutate(sum_prop_wards = scales::percent(round(sum_prop_wards, 3))) |> 
    
    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           sum_prop_wards = if_else(is.na(sum_prop_wards), "0%", sum_prop_wards)) |> 
    select(Hymn = label, `percent of wards` = sum_prop_wards) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Since February 16")
```

I've now seen all the new hymns sung at least once in my dataset. <hymn>Let Easter Anthems Ring (#1205)</hymn> is high on this list, but that's only because of [Easter](/posts/easter2025). It looks like <hymn>Softly and Tenderly Jesus Is Calling (#1020)</hymn> is currently the most popular of the new hymns generally. It's followed by <hymn>This Little Light of Mine (#1028)</hymn> and then <hymn>I Know That the Savior Loves Me (1021)</hymn>, but we'll likely see this change a little bit over the next few weeks and months. 

(For a complete version of this table that includes all hymns, see [here](/overview.qmd).)

## What hymns are most popular each week?

It's been long enough that we can get an idea of which hymns were popular when. The following plot shows the percentage of wards that sang each hymn *of the wards that sang any third batch hymns*. 

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 13}
#| column: page
meetings_since_release3 |> 
    filter(hymn_num %in% c(1019:1031, 1205:1206)) |>
    count(date, hymn_num, name_num) |> 
    
    # pivot and stuff to get zeros
    pivot_wider(names_from = date, values_from = n, values_fill = 0) |> 
    pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
    mutate(date = ymd(date)) |> 
    
    # put them in order
    arrange(hymn_num) |> 
    mutate(name_num = fct_inorder(name_num)) |> 
    
    # Put the dates in order
    arrange(date) |> 
    mutate(date = date) |> 
    
    # proportions per week
    mutate(prop = n/sum(n), .by = date) |> 
    
    ggplot(aes(date, prop, group = name_num)) + 
    geom_path() + 
    geom_point() + 
    scale_x_date(breaks = batch3_date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.1, 0)) +
    scale_y_continuous(labels = scales::percent) + 
    facet_wrap(~name_num, ncol = 4) + 
    labs(title = "Proportion of wards singing new hymns by week",
         subtitle = "Of the sacrament meetings with new hymns",
         x = NULL,
         y = "proportion of wards") + 
    theme_bw()
```

It seems like the most popular hymns are somewhat erratic: <hymn>Softly and Tenderly Jesus Is Calling (#1020)</hymn>, <hymn>I Know That My Savior Loves Me (#1021)</hymn>, and <hymn>This Little Light of Mine (#1028)</hymn> jump around a lot for some reason. It's possibly a sample size issue, because relatively few wards are singing them. Some of the less common ones don't show a clear pattern, other than they're generally present but not common. Some become slightly more popular after a few weeks, again suggesting (like the other batches) that once wards have gotten the favorites out of the way, they are then open to trying new ones. The most obvious trend though is <hymn>Let Easter Anthems Ring (#1205)</hymn> and <hymn>Were You There? (#1206)</hymn> being sung around [Easter](/posts/easter2025).

## When during meetings are these hymns sung?

We can now see when during sacrament meetings these hymns have been sung. These results will likely change slightly as more time passes, but it's been long enough and enough data has come in that we can get a good idea of what's going on.

```{r, include = TRUE}
meetings_since_release3 |> 
    filter(hymn_num %in% c(1019:1031, 1205:1206),
           !is.na(type)) |>
    count(name_num, hymn_num, type) |> 
    mutate(n = round(n/sum(n), 2)*100, 
           .by = name_num) |> 
    pivot_wider(names_from = type, values_from = n, values_fill = 0) |>
    arrange(hymn_num) |> 
    select(-hymn_num) |> 
    select(name_num, Opening, Sacrament, Intermediate, Closing) |> 
    rename(`New Hymn` = name_num) |> 
    gt() |> 
    gt_plt_bar_pct(column = c(Opening), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |> 
    gt_plt_bar_pct(column = c(Sacrament), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Intermediate), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Closing), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Since February 15") |> 
    tab_source_note("Note: Each row adds up to 100%.")
```

The most notable pattern is that <hymn>I Have Faith in the Lord Jesus Christ (#1024)</hymn> is pretty much only sung as a closing hymn. But, it's the least popular hymn in this batch and I've only seen nine wards sing it. Most hymns are sung either as opening or closing hymns, with roughly equal frequencies. Some, like <hymn>Faith in Every Footstep (#1022)</hymn> are sung equally as opening, intermediate, or closing hymns. <hymn>This Little Light of Mine (#1028)</hymn> stands out as being the only one that is most popular as an intermediate hymn. We didn't get any that were obviously intended for the sacrament this time.


## Conclusion

Now that it has been a couple months since the third batch was released, we can see which are more popular than others and when they're sung during the meeting. What is clear though is the relatively cold reception these hymns got in the first month. But, people have warmed up to them, especially the Easter ones. 