---
title: "The Second Batch of New Hymns!"
author: "Joey Stanley"
date: "2024-09-18"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd(floor_date(today(), "week"))
```




```{r}
meetings_since_release1 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-06-02"))
meetings_since_release2 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-09-15"))
n_meetings_since_release1 <- meetings_since_release1 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release2 <- meetings_since_release2 |> 
    distinct(meeting_id) |> 
    nrow()
n_wards_since_release1 <- meetings_since_release1 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release2 <- meetings_since_release2 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_weeks_since_release1 <- meetings_since_release1 |> 
    count(date) |> 
    nrow()
n_weeks_since_release2 <- meetings_since_release2 |> 
    count(date) |> 
    nrow()
```


On September 12th, the church released the second batch of new hymns. In [a previous post](/posts/new_hymns_batch1), I covered in detail as much as I could about the first batch of hymns and how they rolled out between June 2 and September 8th. This page covers period between when the second batch came out and when the third batch will come out, whenever that will be and will update regularly as I collect more data. Currently I have data from `r scales::comma(n_meetings_since_release2)` sacrament meetings since September 15th from `r n_wards_since_release1` wards.

But first, here's how much data I have for each week, just so you have an idea of what I'm working with. For what it's worth, I have more data from some of the past few weeks than any other week in my entire dataset.

```{r, include = TRUE}
meetings_since_release2 |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since September 15th")
```

For some reason I have about half as much data from November 10 as I do the other weeks and I'm not sure why because the way I've been collecting data has been the same.

## How many wards sang new hymns each week?

This plot shows what percentage of wards sing new hymns each week. For context, includes all the data since June 2nd when the first batch came out so we can see how the roll-out of the second batch compares to that of the first batch. I've color-coded it by batch: green is for the first batch, red/pink is for the second, and blue is for both combined. This division is only relevant since September 15th when the second batch came out so prior to then, only the blue lines are used. In all cases, a lighter color indicates wards that sang exactly one new hymn while a darker color is for wards that since two or more in a single meeting.

```{r, fig.height = 4, fig.width = 8}
count_new_hymns <- function(.category_name, .hymns) {
    meetings_since_release1 |> 
        mutate(is_new = hymn_num %in% .hymns) |>
        summarize(n_new = sum(is_new), .by = c(date, ward)) |>
        count(date, n_new) |>
        arrange(date) |>
        mutate(prop_new = n/sum(n), .by = date) |>
        mutate(n_new = as.factor(n_new)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_new, values_from = prop_new)
}
data_to_plot <- bind_rows(
    count_new_hymns("batch1", c(1001:1009, 1201:1204)), 
    count_new_hymns("batch2", 1010:1018), 
    count_new_hymns("new",    c(1001:1018, 1201:1204))) |> 
    pivot_longer(cols = -c(date, category), names_to = "n_new", values_to = "prop") |>
    mutate(group = paste(n_new, category)) |>
    arrange(date) |>
    # don't need zeros from all three
    filter(!(n_new == 0 & category != "new")) |>
    # don't need the 3s because they're so rare
    filter(n_new <= 2) |>
    # don't need batch 2 before Sep 15
    filter(!(category == "batch2" & date < ymd("2024-09-15"))) |>
    # don't need batch 1 before Sep 15 either
    filter(!(category == "batch1" & date < ymd("2024-09-15"))) |>
    filter(!is.na(prop)) |> 
    print()



# Place labels a little above the max from most recent three weeks
# (Will need to raise to four as more dates come in.)
labels <- data_to_plot |> 
    slice_max(order_by = date, n = 3, by = n_new) |> 
    summarize(prop = max(prop), 
              date = max(date),
              .by = n_new) |> 
    mutate(label = case_when(n_new == 0 ~ "no new hymns",
                             n_new == 1 ~ "1 new hymn",
                             n_new >= 2 ~ "2 or more new hymns"))
```


```{r, fig.height = 4, fig.width = 8}
# Get date breaks for this post.
date_minor_breaks <- tibble(weeks_since = 0:52,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= "2024-06-02")
date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 3 == 0)
batch2_date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 2 == 0)
```


```{r, include = TRUE, fig.height = 4, fig.width = 8}
ggplot(data_to_plot, aes(date, prop, group = group, color = group)) +
    # stat_smooth(formula = "y ~ x", method = "loess", color = "gray85", alpha = 0.25) +
    geom_path() +
    geom_point() +
    # geom_text(data = labels, aes(y = prop + 0.05, label = label), hjust = 1) +
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("0 new", 
                                  "1 new", "1 batch1", "1 batch2",
                                  "2 new", "2 batch1", "2 batch2"),
                       labels = c("no new hymns", 
                                 "1 new hymn (total)", "1 new hymn (from batch 1)", "1 new hymn (from batch 2)",
                                 "2 new hymns (total)", "2 new hymns (from batch 1)","2 new hymns (from batch 2)"),
                       values = c("gray66", 
                                  "#a6cee3", "#b2df8a", "#fb9a99",
                                  "#1f78b4", "#33a02c", "#e31a1c")) + 
    theme_minimal() +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since June 2",
         subtitle = paste("Based on data from", n_wards_since_release1, "wards"),
         fill = "date")
    # theme(legend.position = "none")
```

With the [first batch](/posts/new_hymns_batch1) of hymns, the reception was very warm. Pretty consistently over the 15 weeks after the first batch of hymns was released, we saw about 30% of wards singing at least one of the new ones, with a small percentage singing two or more. 

Since the introduction of the second batch, I thought it'd be helpful to split it up into various colors. The green ones represent the number of wards singing hymns from the first batch (#1001--1009), with the lighter color for just one and the darker color for two in a meeting. The red/pink dots represent the second batch of hymns. 

In the first week after the release of the second batch, relatively few wards (around 7%) sang any of them. Meanwhile, the first batch continued as normal. This is about a quarter as many wards as what we saw with the first batch. However, the following Sunday, September 22nd, about 20% of wards sang from the second batch, which is about three times as much as the previous Sunday. This is represented by the pink and red lines going up in the plot. Since General Conference, we're seeing more wards sing from the second batch than the first batch, which is in stark contrast to the lukewarm reception they initially got.

What is perhaps more interesting though is not which batch people sang from, but whether they sang any new hymns at all. Look at the blue lines. Those represent the number of wards singing new hymns total, across both batches. They have gone up over the past few weeks. To accompany that, the gray line, which represents the number of wards that sang none of the new hymns, goes down. On October 20th, almost half of the wards I have data from sang at least one of the new hymns! Since then, more wards than ever (around 7%) have sang two new hymns in one meeting. Since then, it appears that it's stabilizing at around 40% of wards singing at least one new hymn. 

## What hymns are most popular?

The following table shows the hymns from the second batch in order of how popular they are. On the right, I have an approximation of how many wards have sung that hymn since its release. 

```{r, include = TRUE}
freqs |> 
    filter(date >= ymd("2024-09-15"),
           date <= today()) |> 
    
    # Because people are repeating 1001 now, I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |> 
    filter(nth_time == 1) |> 
    
    cleveland_plot(return = "table", hymns_to_plot = 1010:1019) |> 
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    filter(hymn_num %in% 1010:1019) |>
    mutate(percent_wards = scales::percent(round(n_weeks_since_release2 / every_x_weeks, 3))) |> 
    
    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           percent_wards = if_else(is.na(percent_wards), "0%", percent_wards)) |> 
    select(Hymn = label, `percent of wards` = percent_wards) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Since September 15")
```


So, perhaps unsurprisingly, <hymn>Amazing Grace (#1010)</hymn> is far and away the most popular in this batch. A distant second is <hymn>My Shepherd Will Supply My Need (#1014)</hymn>. The next three, <hymn>Holding Hands Around the World (#1011)</hymn>, <hymn>Behold the Wounds in Jesus' Hands (#1016)</hymn>, and <hymn>This Is the Christ (#1017)</hymn> jump around in the ranking from week to week, so the dust hasn't quite settled yet for us to see how common each of them are relative to each other.

(For a complete version of this table that includes all hymns, see [here](/overview.qmd).)

## What hymns are most popular each week?

Now that we're several weeks past the release, we can start to see a glimpse of how the adoption of these new hymns has ebbed and flowed and how each hymn has been introduced. The figure below shows what proportion of wards sang what hymns each week, *given that they sang a new hymn that week*. So, we're only focused on wards that have used these hymns. 

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 8, fig.width = 8}
meetings_since_release2 |> 
    filter(hymn_num %in% 1010:1019) |>  
    count(date, hymn_num, name_num) |> 
    
    # pivot and stuff to get zeros
    pivot_wider(names_from = date, values_from = n, values_fill = 0) |> 
    pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
    mutate(date = ymd(date)) |> 
    
    # put them in order
    arrange(hymn_num) |> 
    mutate(name_num = fct_inorder(name_num)) |> 
    
    # Put the dates in order
    arrange(date) |> 
    mutate(date = date) |> 
    
    # proportions per week
    mutate(prop = n/sum(n), .by = date) |> 
    
    ggplot(aes(date, prop, group = name_num)) + 
    geom_path() + 
    geom_point() + 
    scale_x_date(breaks = batch2_date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(labels = scales::percent) + 
    facet_wrap(~name_num, ncol = 2) + 
    labs(title = "Proportion of wards singing new hymns by week",
         subtitle = "Of the sacrament meetings with new hymns",
         x = "date",
         y = "proportion of wards") + 
    theme_bw()
```

So, <hymn>Amazing Grace (#1010)</hymn> has taken the lion's share of the new hymns. In the first four weeks, if a ward sang any from the second batch, about half of the time that hymn was <hymn>Amazing Grace (#1010)</hymn>. But it has waned a bit since the initial release, and other hymns have gotten more popular. Perhaps now that some wards have gotten it out of the way, they've freed themselves up to sing others. This is similar to [what we saw](/posts/new_hymns_batch1) with <hymn>Come, Thou Fount of Every Blessing (#1001)</hymn>, though in a less extreme way. Mostly recently, we're seeing <hymn>God's Gracious Love (#1013)</hymn> and <hymn>Come, Lord Jesus (#1018)</hymn> spike in popularity.


## When during meetings are these hymns sung?

We can see when these hymns tend to be sung and get an overall look at the popularity of the second batch of hymns. Again, it's based on so little data, so take these results with a grain of salt. 

```{r, include = TRUE}
meetings_since_release2 |> 
    filter(hymn_num %in% 1010:1019) |> 
    count(name_num, hymn_num, type) |> 
    mutate(n = round(n/sum(n), 2), 
           n = scales::percent(n),
           .by = name_num) |> 
    pivot_wider(names_from = type, values_from = n, values_fill = "0%") |> 
    arrange(hymn_num) |> 
    select(-hymn_num) |> 
    select(name_num, Opening, Sacrament, Intermediate, Closing) |> 
    rename(`New Hymn` = name_num) |> 
    gt() |> 
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Since September 15")
```


Looks like <hymn>Amazing Grace (#1010)</hymn> is about split between being an opening hymn or a closing hymn, with the possibility of an intermediate hymn being about half as likely. <hymn>My Shepherd Will Supply My Need (#1014)</hymn> and <hymn>Holding Hands Around the World (#1011)</hymn> are pretty even across all three non-sacrament slots and <hymn>This Is the Christ (#1017)</hymn> is about even across all four slots in a sacrament meeting. 

The other hymns show some patterns but because I've seen them far less often, I have to take these distributions with a grain of salt. Some hymns are most likely to be intermediate hymns, <hymn>God's Gracious Love (#1013)</hymn>, and <hymn>Come, Lord Jesus (#1018)</hymn>. Finally, <hymn>Anytime, Anywhere (#1012)</hymn> and <hymn>Oh, the Deep, Deep Love of Jesus (#1015)</hymn> are more common towards the end of the meeting, while <hymn>Behold the Wounds in Jesus' Hands (#1016)</hymn> is most common at the beginning. 

## Conclusion

At first, people didn't seen as eager to incorporate this second batch of hymns into their sacrament meetings right away. It may be the case that wards had their hymns planned already and it took them few weeks still before incorporating this second batch. It may also be because church musicians need time to prepare. However, the same reasons applied to the first batch and far more people sang new hymns in June 2. So, my guess is that the novelty of singing new hymns has waned a little bit, and there's less of a rush to burn through all of them as there might have been over the summer. 

However, we're now several weeks since the release of the second batch, and it's clear that people are warming up to them. Late October was the peak popularity. Now that we're getting into Thanksgiving (in the US) and Christmas, those holidays will take precedent and we'll see fewer new hymns (except of course the new Christmas ones). In January, once the next batch is scheduled to come out, I think we'll see that third batch becoming more popular than either of the other two for a while.
