---
title: "The Fifth Batch of New Hymns!"
author: "Joey Stanley"
date: "2025-09-18"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd(floor_date(today(), "week"))
library(gtExtras)
```


```{r}
general_conference_dates <- ymd("2024-10-06", "2025-04-06", "2025-10-04", "2025-10-05")
get_meetings_since_release <- function(.date) {
    freqs |> 
        filter(date >= ymd(.date),
               date <= last_sunday,
               !date %in% general_conference_dates)
}
meetings_since_release1 <- get_meetings_since_release("2024-06-02")
meetings_since_release2 <- get_meetings_since_release("2024-09-15")
meetings_since_release3 <- get_meetings_since_release("2025-02-16")
meetings_since_release4 <- get_meetings_since_release("2025-06-15")
meetings_since_release5 <- get_meetings_since_release("2025-09-21")

get_n_meetings_since_release <- function(.df) {nrow(distinct(.df, meeting_id))}
n_meetings_since_release1 <- get_n_meetings_since_release(meetings_since_release1)
n_meetings_since_release2 <- get_n_meetings_since_release(meetings_since_release2)
n_meetings_since_release3 <- get_n_meetings_since_release(meetings_since_release3)
n_meetings_since_release4 <- get_n_meetings_since_release(meetings_since_release4)
n_meetings_since_release5 <- get_n_meetings_since_release(meetings_since_release5)

get_n_wards_since_release <- function(.df) {
    .df |> 
        filter(date <= ymd(last_sunday)) |> 
        pull(ward) |> 
        unique() |> 
        length()
}
n_wards_since_release1 <- get_n_wards_since_release(meetings_since_release1)
n_wards_since_release2 <- get_n_wards_since_release(meetings_since_release2)
n_wards_since_release3 <- get_n_wards_since_release(meetings_since_release3)
n_wards_since_release4 <- get_n_wards_since_release(meetings_since_release4)
n_wards_since_release5 <- get_n_wards_since_release(meetings_since_release5)

get_n_weeks_since_release <- function(.df) {nrow(count(.df, date))}
n_weeks_since_release1 <- get_n_weeks_since_release(meetings_since_release1)
n_weeks_since_release2 <- get_n_weeks_since_release(meetings_since_release2)
n_weeks_since_release3 <- get_n_weeks_since_release(meetings_since_release3)
n_weeks_since_release4 <- get_n_weeks_since_release(meetings_since_release4)
n_weeks_since_release5 <- get_n_weeks_since_release(meetings_since_release5)
```


On September 18, 2025, the church [released](https://newsroom.churchofjesuschrist.org/article/hymn-written-by-president-nelson-included-in-latest-release) the fifth batch of new hymns, which features another African American spiritual, a hymn written by a Tongan saint, and lyrics written by President Nelson. In previous posts, I have covered in detail the [first](/posts/new_hymns_batch1), [second](/posts/new_hymns_batch2),  [third](/posts/new_hymns_batch3), and [fourth](/posts/new_hymns_batch4) batches of hymns and how they were incorporated into sacrament meetings. This page covers period since when the fifth batch came out and will update weekly as more data comes in. 

Currently I have data from `r scales::comma(n_meetings_since_release5)` sacrament meetings from `r scales::comma(n_wards_since_release5)` wards since September 21, 2025. Here's how much data I have for each week so far, just so you have an idea of what I'm working with.

```{r, include = TRUE}
meetings_since_release5 |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since September 21, 2025")
```



## How many wards sang new hymns each week?

The following plot show what percentage of wards sang from each of the new batches of hymns, per week, since June 2024 when the new hymns started getting rolled out. I've included data from all batches just so you can compare them to each other. There's not a lot to see quite yet because it's only been a few weeks, but we an already see some cool trends.

```{r, fig.height = 4, fig.width = 8}
count_new_hymns <- function(.category_name, .hymns) {
    meetings_since_release1 |> 
        mutate(is_new = hymn_num %in% .hymns) |>
        summarize(n_new = sum(is_new), .by = c(date, ward)) |>
        count(date, n_new) |>
        arrange(date) |>
        mutate(prop_new = n/sum(n), .by = date) |>
        mutate(n_new = as.factor(n_new)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_new, values_from = prop_new)
}
data_to_plot <- meetings_since_release1 |> 
    summarize(batch1 = sum(hymn_num %in% batch1_hymn_nums) > 0,
              batch2 = sum(hymn_num %in% batch2_hymn_nums) > 0,
              batch3 = sum(hymn_num %in% batch3_hymn_nums) > 0,
              batch4 = sum(hymn_num %in% batch4_hymn_nums) > 0,
              batch5 = sum(hymn_num %in% batch5_hymn_nums) > 0,
              .by = c(meeting_id, date)) |>
    mutate(none = !batch1 & !batch2 & !batch3 & !batch4 & !batch5) |>
    pivot_longer(cols = c(batch1, batch2, batch3, batch4, batch5, none), names_to = "batch", values_to = "present") |> 
    count(date, batch, present) |> 
    pivot_wider(names_from = present, values_from = n, values_fill = 0) |> 
    mutate(prop_sung = `TRUE`/(`TRUE` + `FALSE`)) |> 
    
    # don't need batch 5 before Sept 21
    filter(!(batch == "batch5" & date < ymd("2025-09-21"))) |>
    # don't need batch 4 before June 15
    filter(!(batch == "batch4" & date < ymd("2025-06-15"))) |>
    # don't need batch 3 before Feb 13
    filter(!(batch == "batch3" & date < ymd("2025-02-13"))) |>
    # don't need batch 2 before Sep 15
    filter(!(batch == "batch2" & date < ymd("2024-09-15"))) |>
    # don't need batch 1 before June 2
    filter(!(batch == "batch1" & date < ymd("2024-06-02"))) |>
    print()
```



```{r, fig.height = 4, fig.width = 8}
weeks_since_initial_release <- ceiling(as.numeric(last_sunday - ymd("2024-06-02")) / 7)

# Get date breaks for this post.
date_minor_breaks <- tibble(weeks_since = 0:weeks_since_initial_release,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= "2024-06-02")
date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 4 == 0)
batch5_date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 1 == 0)
```


```{r}
labels <- tibble(
    batch = c("none", "batch1", "batch2", "batch3", "batch4"),
    prop  = c(0.73, 0.26, 0.08, 0.04, 0.04),
    date  = ymd(c("2025-04-01", "2025-02-14", "2025-01-25", "2025-03-15", "2025-06-15")),
    label = c("no new hymns", "batch 1", "batch 2", "batch 3", "batch4"))

# Annotations
christmas_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2024-12-01",
           date <= "2025-01-01") |> 
    mutate(label = "Christmas")
easter_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2025-04-13",
           date <= "2025-04-27") |> 
    mutate(label = "Easter")
```


```{r, include = TRUE, fig.height = 5.5, fig.width = 14}
#| column: screen
ggplot(data_to_plot, aes(date, prop_sung, group = batch, color = batch)) +
    
    # Christmas annotation
    ggforce::geom_mark_ellipse(data = christmas_data_to_highlight, group = 1, color = "gray50") + 
    annotate(geom = "text", x = ymd("2024-11-03"), y = 0.35, label = "Christmas", hjust = 1, color = "gray50") + 
    annotate(geom = "curve", color = "gray50",
             x = ymd("2024-11-05"), xend = ymd("2024-11-25"), y = 0.35, yend = 0.30,
             curvature = -0.25, arrow = arrow(length = unit(2, "mm"))) + 
    
    # Easter
    ggforce::geom_mark_ellipse(data = easter_data_to_highlight, group = 1, color = "gray50") +
    annotate(geom = "segment", color = "gray50",
             x = ymd("2025-04-20"), xend = ymd("2025-04-20"), y = 0.35, yend = 0.275, 
             arrow = arrow(length = unit(2, "mm"))) +
    annotate(geom = "text", x = ymd("2025-04-20"), y = 0.37, label = "Easter", hjust = 0.5, color = "gray50") +
    
    # Mother's Day
    annotate(geom = "segment", 
             x = ymd("2025-05-11"), xend = ymd("2025-05-11"), y = 0.28, yend = 0.18, 
             arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    annotate(geom = "text", x = ymd("2025-05-11"), y = 0.29, label = "Mother's\nDay", 
             vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +
    
    # Father's Day
    annotate(geom = "segment", 
             x = ymd("2025-06-15"), xend = ymd("2025-06-15"), y = 0.28, yend = 0.18, 
             arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    annotate(geom = "text", x = ymd("2025-06-15"), y = 0.29, label = "Father's\nDay", 
             vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +
    
    
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(data = labels, aes(y = prop, label = label), hjust = 0.5, vjust = 0.5, size = 6) +
    
    
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4", "batch5"),
                       # Color brewer Set1
                       values = c("gray66", "#e41a1c",  "#377eb8",  "#4daf4a", "#984ea3", "#ff7f00")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since June 2",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

The first two weeks of the fifth batch were pretty lukewarm. Very few wards sang from it. A few wards sang <hymn>Our Prayer to Thee (#1048)</hymn> on September 28, [the day after President Nelson died](/posts/president_nelsons_death), but otherwise, fifth-batch hymns were quite rare. 

There is a noticeable jump after General Conference though. More wards sang from the fifth batch on October 12 (7.2%) than even the best week of the fourth batch (August 24th: 6.9%). The bulk of this surge (about 62%) was, perhaps unsurprisingly, <hymn>Our Prayer to Thee (#1048)</hymn>. However, most of the hymns in the fifth batch have been attested at least once in my dataset.


```{r}
# show all fifth batch data
meetings_since_release5 |> 
    filter(hymn_num %in% batch5_hymn_nums)

# by hymn
meetings_since_release5 |> 
    filter(hymn_num %in% batch5_hymn_nums) |> 
    count(hymn_name, hymn_num, sort = TRUE) |> 
    mutate(prop = n/sum(n))
# by hymn and by date
meetings_since_release5 |> 
    filter(hymn_num %in% batch5_hymn_nums) |> 
    count(hymn_name, hymn_num, date, sort = TRUE) |> 
    mutate(prop = n/sum(n), .by = date)

# by date
data_to_plot |> 
    filter(batch == "batch5")
# sorted by best weeks
data_to_plot |> 
    filter(batch == "batch5") |> 
    arrange(-prop_sung)
# fourth batch for comparison
data_to_plot |> 
    filter(batch == "batch4") |> 
    arrange(-prop_sung)

```

Here's another version of the same plot but instead of the actual dates, it's in terms of how many weeks since the batches came out. This version of the plot is useful to spot any parallel trends across the batches, especially in the first few weeks of their release.

```{r, include = TRUE, fig.height = 4.5, fig.width = 14}
#| column: page
data_to_plot |> 
    mutate(weeks_since_release = as.numeric(date - min(date))/7, .by = batch) |> 
    filter(!batch %in% "none") |> 
    ggplot(aes(weeks_since_release, prop_sung, group = batch, color = batch)) +
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    scale_x_continuous(breaks = seq(2, 100, 2),
                       expand = expansion(0, c(0.2, 2))) + 
    scale_y_continuous(expand = expansion(0, c(0.01, 0.02)),
                       breaks = seq(0, 1, 0.1),
                       minor_breaks = seq(0, 1, 0.02),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4", "batch5"),
                       # Color brewer Set 1
                       values = c("gray66", "#e41a1c",  "#377eb8",  "#4daf4a", "#984ea3", "#ff7f00")) +
    labs(x = "weeks since release",
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since their release date",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

In this version of the plot, It's clearer that the fifth batch is doing better than the fourth batch in all weeks so far. It's not quite as good as the third batch though (and certainly not the second or first). 


## What hymns are most popular?

The following table shows the hymns from the fifth batch in order of how popular they are.  

```{r, include = TRUE}
meetings_since_release5 |> 
    
    # I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |> 
    filter(nth_time == 1) |> 
    
    cleveland(return = "table", 
              stat_to_plot = sum_prop_wards,
              hymns_to_plot = batch5_hymn_nums) |> 
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    filter(hymn_num %in% batch5_hymn_nums) |>
    mutate(sum_prop_wards = scales::percent(round(sum_prop_wards, 3))) |> 
    
    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           sum_prop_wards = if_else(is.na(sum_prop_wards), "0%", sum_prop_wards)) |> 
    select(Hymn = label, `percent of wards` = sum_prop_wards) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Since June 15, 2025")
```


It's hard to say whether <hymn>Our Prayer to Thee (#1048)</hymn> would have become as popular as it currently is had President Nelson not passed away recently. We'll never know. But the point is, that hymn is currently the leader in this fifth batch and will likely stay that way until Christmastime when <hymn>Go Tell It on the Mountain (#1208)</hymn> will likly take the lead.



## What hymns are most popular each week?

It's still a little premature to include this plot, but I will anyway. This shows how popular each hymn is each week. Out of all the fifth batch hymns sing in a given week, what percentage of them were <hymn>Our Prayer to Thee (#1048)</hymn>, for example?

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 13}
#| column: page
meetings_since_release5 |>
     filter(hymn_num %in% batch5_hymn_nums) |>
     count(date, hymn_num, name_num) |>

     # pivot and stuff to get zeros
     pivot_wider(names_from = date, values_from = n, values_fill = 0) |>
     pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
     mutate(date = ymd(date)) |>

     # put them in order
     arrange(hymn_num) |>
     mutate(name_num = fct_inorder(name_num)) |>

     # Put the dates in order
     arrange(date) |>
     mutate(date = date) |>

     # proportions per week
     mutate(prop = n/sum(n), .by = date) |>

     ggplot(aes(date, prop, group = name_num)) +
     geom_path() +
     geom_point() +
     scale_x_date(date_breaks = "3 weeks",
                  date_minor_breaks = "1 week",
                  date_labels = "%b %d",
                  expand = expansion(0.1, 0)) +
     scale_y_continuous(labels = scales::percent) +
     facet_wrap(~name_num, ncol = 4) +
     labs(title = "Proportion of wards singing new hymns by week",
          subtitle = "Of the sacrament meetings with new hymns",
          x = NULL,
          y = "proportion of wards") +
     theme_bw()
```

Since the data is still so sparse, it's important to to put too much weight into the percentages. But it is clear that <hymn>Our Prayer to Thee (#1048)</hymn> is the lion's share of fifth batch hymns. Something we're seeing though is that two other hymns, <hymn>Thou Gracious Lord, Whose Mercy Lends (#1042)</hymn> and <hymn>Help Us Remember (#1043)</hymn> have become more common recently, with roughly a quarter to a third of wards that have sung fifth-batch hymns singing them. 


## When during meetings are these hymns sung?

We can also see when during sacrament meetings these hymns have been sung. Data sparsity makes this a pretty unreliable table though. We'll have to wait a few more weeks or months to really see some more meaningful trends. The only one that is somewhat trustworthy is, you guessed it, <hymn>Our Prayer to Thee (#1048)</hymn>.

```{r, include = TRUE}
meetings_since_release5 |>
    filter(hymn_num %in% batch5_hymn_nums,
           !is.na(type)) |>
    count(name_num, hymn_num, type) |>
    mutate(n = round(n/sum(n), 2)*100,
           .by = name_num) |>
    pivot_wider(names_from = type, values_from = n, values_fill = 0) |>
    arrange(hymn_num) |>
    select(-hymn_num) |>
    select(name_num,
           Opening, Sacrament, Intermediate, Closing) |>
    rename(`New Hymn` = name_num) |>
    gt() |>
    gt_plt_bar_pct(column = c(Opening), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Sacrament), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Intermediate), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Closing), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Since September 21, 2025") |>
    tab_source_note("Note: Each row adds up to 100%.")
```



## Conclusion

That's all for now. It's too soon to say anything else because so few wards have sung from the fifth batch. I can't wait to see what happens in the next few weeks though. 
