---
title: "The First Batch of New Hymns!"
author: "Joey Stanley"
date: "2024-06-25"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")
library(gt)

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd(floor_date(today(), "week"))
```




```{r}
meetings_since_release <- freqs |> 
    filter(date <= ymd("2024-09-08"),
           date >= ymd("2024-06-02"))
n_meetings_since_release <- meetings_since_release |> 
    distinct(meeting_id) |> 
    nrow()
n_wards_since_release <- meetings_since_release |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_weeks_since_release <- meetings_since_release |> 
    count(date) |> 
    nrow()
```


On May 30th, the church released the first batch of new hymns. I have done posts on how those hymns have been used during the [first](posts/new_hymns_week1), [second](posts/new_hymns_week2), and [third](posts/new_hymns_week3) weeks since their release, but this page covers the 15 weeks between when the first and second batches came out. When I get more data from this period I will update this page, but it will largely remain as is. Currently I have data from `r scales::comma(n_meetings_since_release)` sacrament meetings between June 2nd and September 8th from `r n_wards_since_release` wards. For now, I'll just focus on hymns 1001--1009 and will wait to bring in the new holiday hymns later on.

But first, here's how much data I have for each week, just so you have an idea of what I'm working with.

```{r, include = TRUE}
meetings_since_release |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since June 2")
```




## How many wards are singing new hymns each week?

The following plot shows the number of new hymns sung per ward each week between the first and second batches of the new hymns.

```{r, include = TRUE, fig.height = 4, fig.width = 8}
data_to_plot <- meetings_since_release |> 
    mutate(is_new = hymn_num > 1000 & hymn_num < 2000) |> 
    summarize(n_new = sum(is_new), .by = c(date, ward)) |>
    count(date, n_new) |> 
    filter(n_new <= 2) |> 
    
    # Put the dates in order
    arrange(date) |> 
    # mutate(date = fct_inorder(sf(date))) |> 
    
    mutate(prop = n/sum(n), .by = date) |>
    mutate(n_new = as.factor(n_new))

# Place labels a little above the max from most recent three weeks
# (Will need to raise to four as more dates come in.)
labels <- data_to_plot |> 
    slice_max(order_by = date, n = 3, by = n_new) |> 
    summarize(prop = max(prop), 
              date = max(date),
              .by = n_new) |> 
    mutate(label = case_when(n_new == 0 ~ "no new hymns",
                             n_new == 1 ~ "1 new hymn",
                             n_new == 2 ~ "2 new hymns"))

date_breaks <- data_to_plot |> 
    distinct(date) |> 
    rowid_to_column("week_num") |> 
    filter(week_num %% 2 == 0) |> 
    pull(date)

ggplot(data_to_plot, aes(date, prop, group = n_new, color = n_new)) +
    stat_smooth(formula = "y ~ x", method = "loess", color = "gray85", alpha = 0.25) +
    geom_path() + 
    geom_point() + 
    geom_text(data = labels, aes(y = prop + 0.05, label = label), hjust = 1) + 
    scale_x_date(breaks = date_breaks,
                 date_labels = "%b %d",
                 expand = expansion(0.01, 0)) +
    scale_y_continuous(expand = expansion(0, 0.01),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(values = c(ptol_blue, ptol_red, ptol_green)) + 
    theme_minimal() +
    labs(x = "number of new hymns per sacrament meeting",
         y = "percentage of wards",
         title = "Percentage of wards singing new hymns between June 2 and Sept 8",
         subtitle = paste("Based on data from", n_wards_since_release, "wards"),
         fill = "date") + 
    theme(legend.position = "none")
```

During this period, the number of wards that have sung new hymns has been remarkably stable. Typically, only about a quarter to a third of wards sing new hymns each week. A small number sing two each week. There are rare cases of three or more, but they are so infrequent, I've left them off this chart. 

For each number of new hymns per week, I've overlaid a smooth to show the general trend. What I like about it is that is gives a shaded area showing where the proportions typically occur. Anything outside of those could be considered exceptional in some way. In this period here is one week that kinda stands out: August 4th. For some reason, that was the week where the fewest wards sang new hymns. I have no explanation for why that would be. Even when competing with holidays like [Father's Day](/posts/fathers_day), the [martyrdom of Joseph Smith](/posts/martyrdom), and [Pioneer Day](/posts/pioneer_day), the number of new hymns per week stayed consistent. But a random week in early August had fewer---albeit by a small margin. 

## How many new hymns do wards typically sing?

Another way of looking at the data is to consider how many new hymns a ward typically sang in the 15 weeks between the first and second batches. Did a lot of wards sing one a week? Just one the entire time? What is the typical number of new hymns a ward has sung during this period?

It's hard to say for sure with the data I have because most of it comes from scattered contributions from various wards. Many wards in this sample contributed data from multiple weeks, but might have missed one or two because of stake conference or whatever other reason. So, if I've only got ten weeks of data from a particular ward during this 15 week period, and in none of those 10 did they sing new hymns, it's impossible to say whether they sang none at all or because I just happened to not get data from the weeks that they did sing new hymns.

```{r}
n_weeks_since_release
wards_with_complete_data <- meetings_since_release |> 
    summarize(n_new = sum(hymn_num > 1000 & hymn_num < 2000), .by = c(ward, date)) |> 
    arrange(ward, date) |> 
    
    # only wards with (nearly) complete datasets
    add_count(ward, name = "n_weeks") |> 
    filter(n_weeks >= n_weeks_since_release * 0.8) |> 
    
    summarize(weeks_with_new = sum(n_new >= 1), .by = ward) |> 
    count(weeks_with_new) |> 
    mutate(prop = n/sum(n)) |> 
    print()
n_wards_with_complete_data <- meetings_since_release |> 
    distinct(date, ward) |> 
    count(ward) |> 
    filter(n == 2) |> 
    nrow() |> 
    print() 
ggplot(wards_with_complete_data, aes(weeks_with_new, prop)) + 
    geom_col() + 
    scale_x_continuous(name = "weeks with a new hymn",
                       breaks = 0:100) + 
    scale_y_continuous(labels = scales::percent, 
                       name = "percent of wards", 
                       expand = expansion(0, 0)) + 
    labs(title = "How many weeks since June 2 have wards included new hymns?",
         subtitle = paste("Based on the", n_wards_with_complete_data, "wards with >90% complete data")) + 
    theme_minimal()
```

I took a look at the `r n_wards_with_complete_data` wards that I have data from from at least 90% of the 15 weeks in this period. About 15 of those wards didn't sing any new hymns, with the most number of wards sing five. But, that's only based on `r n_wards_with_complete_data` wards and not the complete dataset, so it might not be the best bit of information.


So, I decided to look at all the data I have and organize it a little differently. For each ward, I counted how many weeks' worth of data I have as well as how many new hymns they sang. The following plot shows this data organized kinda like a table. It's a little hard to interpret at first, so let me explain (and you may want to click on the "plot with annotations tab" to help with interpretation). The numbers in each column add up to 100%. So, to go back to the previous paragraph, of the wards I have fifteen weeks of data, 23% of them sang five new hymns. Of the wards I have nine weeks of data from, 43% of them sang just one new hymn. Across the plot, higher numbers are accompanied by darker rectangles.

::: {.panel-tabset}

```{r}
news_per_ward <-  meetings_since_release |> 
    summarize(n_weeks = length(unique(date)),
              n_new = sum(hymn_num %in% 1001:1009),
              .by = "ward") |>
    count(n_weeks, n_new) |> 
    mutate(prop_per_n = n/sum(n), .by = n_weeks) |> 
    mutate(prop_label = paste0(round(prop_per_n*100), "%")) |> 
    filter(n_new < 19)

p <- ggplot(news_per_ward, aes(n_weeks, n_new, color = prop_per_n)) + 
    geom_point(shape = 15, size = 9, position = position_nudge(x = -0.1)) + 
    geom_point(shape = 15, size = 9, position = position_nudge(x =  0.1)) +
    geom_text(aes(label = prop_label), color = "white") + 
    scale_x_continuous(breaks = 1:15,
                       expand = expansion(0, 0.5),
                       limits = c(0.5, 15.5)) + 
    scale_y_continuous(breaks = 0:20,
                       minor_breaks = NULL,
                       limits = c(0, 13),
                       expand = expansion(0, 0.5)) + 
    scale_color_distiller(palette = "Purples", direction = 1) + 
    labs(title = "How many new hymns did wards sing over 15 weeks?",
         subtitle = "Percents refer to number of wards per column",
         x = "number of weeks I have data from per ward",
         y = "number of new hymns") + 
    theme_minimal() + 
    theme(legend.position = "none")
```

### Plain plot

```{r, include = TRUE, fig.height = 5, fig.width = 8}
p
```

### Plot with annotations

```{r, include = TRUE, fig.height = 5, fig.width = 8}
p + 
    
    # Arrow pointing to the column
    annotate(geom = "curve", xend = 9, x = 8, yend = 8, y = 10, curvature = -0.3,
             arrow = arrow(angle = 20, length = unit(0.2, "cm"), type = "closed")) + 
    annotate(geom = "segment", x = 8.5, xend = 9.5, y = 7.8, yend = 7.8) + 
    annotate(geom = "segment", x = 8.5, xend = 8.5, y = 7.8, yend = 7.5) + 
    annotate(geom = "segment", x = 9.5, xend = 9.5, y = 7.8, yend = 7.5) + 
    annotate(geom = "text", x = 8, y = 10, label = "cells in each column\nadd up to 100%",
             hjust = 1) + 
    
    # highlighting the nine box
    annotate(geom = "rect", xmin = 8.5, xmax = 9.5, ymin = 0.5, ymax = 1.5, 
             fill = "#FFFFFF00", color = "#f1a340", linewidth = 1) + 
    annotate(geom = "curve", xend = 8.75, x = 7, yend = 1.5, y = 6.5, curvature = -0.2,
             arrow = arrow(angle = 20, length = unit(0.2, "cm"), type = "closed")) +
    annotate(geom = "text", x = 7, y = 7, label = "Of the wards I have have\nnine weeks of data from, 33% of\nthem sang just 1 new hymn.",
             hjust = 1) + 
    
    # highlighting the 15 box
    annotate(geom = "rect", xmin = 14.5, xmax = 15.5, ymin = 4.5, ymax = 5.5, 
             fill = "#FFFFFF00", color = "#f1a340", linewidth = 1) + 
    annotate(geom = "curve", xend = 14.75, x = 13, yend = 5.5, y = 12, curvature = -0.2,
             arrow = arrow(angle = 20, length = unit(0.2, "cm"), type = "closed")) +
    annotate(geom = "text", x = 13, y = 12.5, label = "Assuming a full dataset of\n15 weeks, the most typical\nnumber of new hymns was 5.",
             hjust = 1)
```


### Plot with line

```{r}
summary(lm(n_new ~ n_weeks, data = news_per_ward))
```
```{r, include = TRUE, fig.height = 5, fig.width = 8}
p + 
    stat_smooth(method = "lm", formula = "y ~ x", color = "#3f007d")
```


:::

So, in this plot, there is a general trend upward: the more data I have from a ward (so, the further left go you in the plot), the more new hymns I see in their data (the darker cells are higher up). I've illustrated this by overlaying a line that tries to capture this trend as best as possible (which you can see by clicking on the "plot with line tab").[For the statisticians there, it's just a simple linear regression line with an intercept of 0.21, a slope of 0.333, and a *p*-value < 0.001. The adjusted r-squared is 0.21]{.aside} The takeaway here is that the line suggests that the typical ward introduces a new hymn about once every three weeks. 

So, while this technically shows the number of new hymns I have from each ward given the amount of data I have in that ward, I think it's safe to interpret as the number of hymns wards are likely to sing after that many weeks since the release of the new hymns. So, after nine weeks (so, by July 28th), we can estimate that about 43% of wards have sung just one new hymn. This is a safe interpretation as long as the rate at which wards introduce new hymns is consistent during these 15 weeks (and judging from the plot in the previous section, that might be true) and if there's not a systemtic bias against *not* getting data when a ward is more or less likely to sing a new hymn (which I can't imagine being the case). It's not a perfect interpretation, but it's another look at the data.


## What hymns are most popular?

It should come as no surprise that the most popular hymn in this new batch is <hymn>Come, Thou Fount of Every Blessing (#1001)</hymn>. But just how popular was it? The following table shows the new hymns in order of how often they were sung. The percent column shows what percentage of wards each hymn has been sung in as congregational hymns during this period.[Note that these numbers might be slightly inflated because some wards have sung the more popular hymns more than once.]{.aside}


```{r, include = TRUE}
meetings_since_release |> 
    
    # Because people are repeating 1001 now, I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |> 
    filter(nth_time == 1) |> 
    
    cleveland_plot(return = "table") |> 
    
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    
    filter(hymn_num %in% c(1001:1009)) |> 
    
    mutate(percent_wards = scales::percent(round(n_weeks_since_release / every_x_weeks, 3))) |> 
    select(Hymn = label, `percent of wards` = percent_wards) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Between June 2 and September 8")
```

Overall then, it seems like the hymns have been pretty well-incorporated into regular singing. An overwhelming majority of wards have sung <hymn>Come, Thou Fount of Every Blessing (#1001)</hymn> between June 2 and September 8, which is pretty remarkable. Surely the novelty will wear off a little bit, but if that pace continues, then wards will sing it 3--4 times a year, or once every 13 weeks. For reference, before June 2nd, the most common hymn was <hymn>I Stand All Amazed (#193)</hymn> and was sung about once every 19 weeks per ward. The other hymns have been adopted with varying amounts of enthusiasm. We'll wait until December to see how the new Christmas ones get used. 

(For a complete version of this table that includes all hymns, see [here](/overview.qmd).)

## What hymns are most popular each week?

We can see to see a glimpse of how the adoption of these new hymns has ebbed and flowed and how each hymn has been introduced. The figure below shows what proportion of wards sang what hymns each week, *given that they sang a new hymn that week*. So, we're only focused on wards that have used these hymns. 

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 8, fig.width = 8}
meetings_since_release |> 
    filter(hymn_num %in% 1001:1009) |>  
    count(date, hymn_num, name_num) |> 
    
    # pivot and stuff to get zeros
    pivot_wider(names_from = date, values_from = n, values_fill = 0) |> 
    pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
    mutate(date = ymd(date)) |> 
    
    # put them in order
    arrange(hymn_num) |> 
    mutate(name_num = fct_inorder(name_num)) |> 
    
    # Put the dates in order
    arrange(date) |> 
    mutate(date = date) |> 
    
    # proportions per week
    mutate(prop = n/sum(n), .by = date) |> 
    
    ggplot(aes(date, prop, group = name_num)) + 
    geom_path() + 
    geom_point() + 
    scale_x_date(date_breaks = "2 weeks", date_labels = "%m/%d") +
    scale_y_continuous(labels = scales::percent) + 
    facet_wrap(~name_num, ncol = 2) + 
    labs(title = "Proportion of wards singing new hymns by week",
         subtitle = "Of the sacrament meetings with new hymns",
         x = "date",
         y = "proportion of wards") + 
    theme_bw()

```

<hymn>Come, Thou Fount of Every Blessing (#1001)</hymn> was most popular during the first week. Of the wards that sang new hymns, two thirds of them sang it on June 2nd. My guess is we will never see a hymn hit that kind of popularity again, with the possible exception of new holiday hymns. Since then, the popularity of <hymn>Come, Thou Fount of Every Blessing (#1001)</hymn> has gone down. It went through a very tidy rise and fall, and then suddenly on August 25, it saw a huge spike again. Perhaps a lot of wards are singing it a second time now.

There are a few other major trends so far. <hymn>As Bread is Broken (#1007)</hymn> was very popular on June 16th, Father's day. A common trend was to sing Father's Day songs for opening and closing hymns and that hymn as the sacrament hymn. Also, <hymn>It Is Well With My Soul (#1003)</hymn> got a little more popular July 14th and has been more popular since then. <hymn>His Eye Is on the Sparrow (#1005)</hymn> saw a spike on July 28th and appears to be more popular since then. Finally, <hymn>I Will Walk with Jesus (#1004)</hymn> wasn't very popular during the first two weeks, but it has been since then and was most popular on August 11th.

## When during meetings are these hymns sung?

We can see when these hymns tend to be sung and get an overall look at the popularity of the first batch of hymns. 

```{r, include = TRUE}
meetings_since_release |> 
    filter(hymn_num %in% 1001:1009) |> 
    count(name_num, hymn_num, type) |> 
    mutate(n = round(n/sum(n), 2), 
           n = scales::percent(n),
           .by = name_num) |> 
    pivot_wider(names_from = type, values_from = n, values_fill = "0%") |> 
    arrange(hymn_num) |> 
    select(-hymn_num) |> 
    select(name_num, Opening, Sacrament, Intermediate, Closing) |> 
    rename(`New Hymn` = name_num) |> 
    gt() |> 
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Between June 2 and September 8")
```

Each hymn has its own story. Unsurprisingly, the hymns that are most likely to be sacrament hymns are <hymn>As Bread is Broken (#1007)</hymn>, <hymn>Bread of Life, Living Water (#1008)</hymn>, and <hymn>Gethsemane (#1009)</hymn>, with that last one being sung pretty often elsewhere in the meeting. <hymn>Think a Sacred Song (#1006)</hymn> appears to be a little more common in the second half of the meeting, but I have only seen it sung in a few wards so that may change if more data comes in. <hymn>Come, Thou Fount of Every Blessing (#1001)</hymn> is about as likely to be an opening hymn as a closing hymn. <hymn>It Is Well with My Soul (#1003)</hymn> is most common as a clsoing hymn. <hymn>His Eye Is on the Sparrow (#1005)</hymn> is about evently split between all three slots. The last two hymns, <hymn>I Will Walk with Jesus (#1004)</hymn> and <hymn>When the Savior Comes Again (#1002)</hymn> are a little more commonly sung as intermediate hymns.


## Conclusion

This post shows a handful of trends in the first batch of hymns in the first 15 weeks since their release. Again, this is only based on `r n_meetings_since_release` sacrament meetings from `r n_wards_since_release` wards. It's not an insignificant amount of data, but it is just a sliver of the many wards across the country and the world. As always, I'm collecting more data and if you'd like to contribute your ward's data, you can do so [here](/contribute.html).


