---
title: "The Fourth Batch of New Hymns!"
author: "Joey Stanley"
date: "2025-06-15"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd("2025-09-13")
library(gtExtras)
```



```{r}
meetings_since_release1 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-06-02"),
           !date %in% ymd("2024-10-06",
                          "2025-04-06"))
meetings_since_release2 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-09-15"),
           !date %in% ymd("2024-10-06",
                        "2025-04-06"))
meetings_since_release3 <- freqs |> 
    filter(date <= ymd("2025-06-08"),
           date >= ymd("2025-02-16"),
           !date %in% ymd("2025-04-06"))
meetings_since_release4 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2025-06-15"),
           date <= ymd("2025-09-13"))

n_meetings_since_release1 <- meetings_since_release1 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release2 <- meetings_since_release2 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release3 <- meetings_since_release3 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release4 <- meetings_since_release4 |> 
    distinct(meeting_id) |> 
    nrow()

n_wards_since_release1 <- meetings_since_release1 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release2 <- meetings_since_release2 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release3 <- meetings_since_release3 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release4 <- meetings_since_release4 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()

n_weeks_since_release1 <- meetings_since_release1 |> 
    count(date) |> 
    nrow()
n_weeks_since_release2 <- meetings_since_release2 |> 
    count(date) |> 
    nrow()
n_weeks_since_release3 <- meetings_since_release3 |> 
    count(date) |> 
    nrow()
n_weeks_since_release4 <- meetings_since_release4 |> 
    count(date) |> 
    nrow()
```



On June 12th, 2025, the church [released](https://newsroom.churchofjesuschrist.org/article/eleven-new-hymns-now-available-in-hymns-for-home-and-church) the fourth batch of new hymns. In previous posts, I have covered in detail the [first](/posts/new_hymns_batch1), [second](/posts/new_hymns_batch2), and [third](/posts/new_hymns_batch3) batches of hymns and how they were incorporated into sacrament meetings. This page covers period since when the fourth batch came out through September 13th, the week before the fifth batch came out.

Currently I have data from `r scales::comma(n_meetings_since_release4)` sacrament meetings from `r scales::comma(n_wards_since_release4)` wards since June 15, 2025. Here's how much data I have for each week during this period, just so you have an idea of what I'm working with.

```{r, include = TRUE}
meetings_since_release4 |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since June 15th, 2025")
```



## How many wards sang new hymns each week?

The following plot show what percentage of wards sang from each of the new batches of hymns, per week, since June 2024 when the new hymns started getting rolled out. I've included data from all batches just so you can compare them to each other.

```{r, fig.height = 4, fig.width = 8}
count_new_hymns <- function(.category_name, .hymns) {
    meetings_since_release1 |> 
        mutate(is_new = hymn_num %in% .hymns) |>
        summarize(n_new = sum(is_new), .by = c(date, ward)) |>
        count(date, n_new) |>
        arrange(date) |>
        mutate(prop_new = n/sum(n), .by = date) |>
        mutate(n_new = as.factor(n_new)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_new, values_from = prop_new)
}
data_to_plot <- meetings_since_release1 |> 
    summarize(batch1 = sum(hymn_num %in% c(1001:1009, 1201:1204)) > 0,
              batch2 = sum(hymn_num %in% c(1010:1018)) > 0,
              batch3 = sum(hymn_num %in% c(1019:1031, 1205:1206)) > 0,
              batch4 = sum(hymn_num %in% c(1032:1041, 1207)) > 0,
              .by = c(meeting_id, date)) |>
    mutate(none = !batch1 & !batch2 & !batch3 & !batch4) |>
    pivot_longer(cols = c(batch1, batch2, batch3, batch4, none), names_to = "batch", values_to = "present") |> 
    count(date, batch, present) |> 
    pivot_wider(names_from = present, values_from = n, values_fill = 0) |> 
    mutate(prop_sung = `TRUE`/(`TRUE` + `FALSE`)) |> 
    
    filter(!(batch == "batch4" & date < ymd("2025-06-15"))) |>
    # don't need batch 3 before June 15
    filter(!(batch == "batch3" & date < ymd("2025-02-13"))) |>
    # don't need batch 2 before Sep 15
    filter(!(batch == "batch2" & date < ymd("2024-09-15"))) |>
    # don't need batch 1 before Sep 15 either
    filter(!(batch == "batch1" & date < ymd("2024-06-02"))) |>
    print()
```



```{r, fig.height = 4, fig.width = 8}
weeks_since_initial_release <- ceiling(as.numeric(last_sunday - ymd("2024-06-02")) / 7)

# Get date breaks for this post.
date_minor_breaks <- tibble(weeks_since = 0:weeks_since_initial_release,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= "2024-06-02")
date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 4 == 1)
batch4_date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 1 == 0)
```


```{r}
labels <- tibble(
    batch = c("none", "batch1", "batch2", "batch3", "batch4"),
    prop  = c(0.73, 0.26, 0.08, 0.04, 0.04),
    date  = ymd(c("2025-04-01", "2025-02-14", "2025-01-25", "2025-03-15", "2025-06-15")),
    label = c("no new hymns", "batch 1", "batch 2", "batch 3", "batch4"))

# Annotations
christmas_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2024-12-01",
           date <= "2025-01-01") |> 
    mutate(label = "Christmas")
easter_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2025-04-13",
           date <= "2025-04-27") |> 
    mutate(label = "Easter")
```


```{r, include = TRUE, fig.height = 5.5, fig.width = 14}
#| column: page
ggplot(data_to_plot, aes(date, prop_sung, group = batch, color = batch)) + 

    geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray80") +

    # Christmas annotation
    ggforce::geom_mark_ellipse(data = christmas_data_to_highlight, group = 1, color = "gray50") +
    annotate(geom = "text", x = ymd("2024-11-03"), y = 0.35, label = "Christmas", hjust = 1, color = "gray50") +
    annotate(geom = "curve", color = "gray50",
             x = ymd("2024-11-05"), xend = ymd("2024-11-25"), y = 0.35, yend = 0.30,
             curvature = -0.25, arrow = arrow(length = unit(2, "mm"))) +

    # Easter
    ggforce::geom_mark_ellipse(data = easter_data_to_highlight, group = 1, color = "gray50") +
    annotate(geom = "segment", color = "gray50",
             x = ymd("2025-04-20"), xend = ymd("2025-04-20"), y = 0.35, yend = 0.275,
             arrow = arrow(length = unit(2, "mm"))) +
    annotate(geom = "text", x = ymd("2025-04-20"), y = 0.37, label = "Easter", hjust = 0.5, color = "gray50") +

    # Mother's Day
    annotate(geom = "segment",
             x = ymd("2025-05-11"), xend = ymd("2025-05-11"), y = 0.28, yend = 0.18,
             arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    annotate(geom = "text", x = ymd("2025-05-11"), y = 0.29, label = "Mother's\nDay",
             vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +

    # Father's Day
    annotate(geom = "segment",
             x = ymd("2025-06-15"), xend = ymd("2025-06-15"), y = 0.28, yend = 0.18,
             arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    annotate(geom = "text", x = ymd("2025-06-15"), y = 0.29, label = "Father's\nDay",
             vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +


    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(data = labels, aes(y = prop, label = label), hjust = 0.5, vjust = 0.5, size = 6) +


    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4"),
                       # Color brewer Set1
                       values = c("gray66", "#e41a1c",  "#377eb8",  "#4daf4a", "#984ea3")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since June 2",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), 
                          "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

```{r}
as_percent_joey <- function(x){
    paste0(round(x * 100, 1), "%")
}
```

During this period, the fourth batch was never very popular. August 24th was the best week with 6.7% of wards singing from it. There was a little bit of jumping around with the second batch, but the general trend is that the first batch was the most common, then the second, then the third, then the fourth. 

Interestingly, we are now to the point where about half of wards on a given Sunday sing at least one new hymn. We saw that for the first time on May 24th due to a spike in the third batch that day. (No one hymn was responsible for that spike, so I think it was just a fluke.) But, for three weeks in a row in August, we saw more than half of wards singing new hymns. I think thit is an important milestone and shows just how much these new hymns are being used in church.

```{r}
data_to_plot |> 
    filter(batch == "batch4")
data_to_plot |> 
    filter(batch == "batch4") |> 
    arrange(-prop_sung)
```

Here's another version of the same plot but instead of the actual dates, it's in terms of how many weeks since the batches came out. This version of the plot is useful to spot any parallel trends across the batches, especially in the first few weeks of their release.

```{r, include = TRUE, fig.height = 4.5, fig.width = 14}
#| column: page
data_to_plot |> 
    mutate(weeks_since_release = as.numeric(date - min(date))/7, .by = batch) |> 
    filter(!batch %in% "none") |> 
    ggplot(aes(weeks_since_release, prop_sung, group = batch, color = batch)) +
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    scale_x_continuous(breaks = seq(2, 100, 2),
                       expand = expansion(0, c(0.2, 2))) + 
    scale_y_continuous(expand = expansion(0, c(0.01, 0.02)),
                       breaks = seq(0, 1, 0.1),
                       minor_breaks = seq(0, 1, 0.02),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4"),
                       # Color brewer Set 1
                       values = c("gray66", "#e41a1c",  "#377eb8",  "#4daf4a", "#984ea3")) +
    labs(x = "weeks since release",
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since their release date",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), 
                          "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

```{r}
n_wards_june15 <- meetings_since_release4 |>
    filter(date == min(date)) |>
    distinct(meeting_id) |>
    nrow() |>
    print()
```


Here we can see that each batch is a little less popular than the previous one, both overall and week by week. Nearly 20% of wards sang from the first batch the week it came out. About 8% did for the second batch. About 1% for the third batch. And of the `r n_wards_june15` wards I have data from on June 15th, just one of them sang from the fourth batch. Granted, it was Father's Day, after several weeks, still very few wards had given the hymns in this new batch a try.


## Multiple new hymns

So, rather than splitting up the data by batch, let's look at just how many new hymns wards sing each week. In this plot, I show the percentage of wards that sing zero, one, or at least two hymns per week.


```{r}
n_new_by_week012 <- meetings_since_release1 |> 
    mutate(is_new = hymn_num > 1000) |> 
    summarize(n_new = sum(is_new), .by = c(date, meeting_id)) |> 
    mutate(n_new = as.factor(if_else(n_new >= 2, "2+", as.character(n_new)))) |> 
    count(date, n_new) |>
    mutate(prop = n/sum(n), .by = date) |> 
    print()
```


```{r}
n_per_week_labels012 <- tibble(date = ymd(c("2025-01-10", "2024-12-15", "2024-11-20")),
                            prop = c(0.77, 0.45, 0.11),
                            n_new = c("0", "1", "2+"),
                            label = c("no new hymns", "1 new hymn", "2 or more new hymns"))
n_per_week_batch_labels <- tibble(date = ymd(c("2024-09-15", "2025-02-16", "2025-06-15")),
                                  prop = c(0.9),
                                  label = c("batch 2", "batch 3", "batch 4"))
```


```{r, include = TRUE, fig.height = 4.5, fig.width = 14}
#| column: page
#| 
ggplot(n_new_by_week012, aes(date, prop, color = n_new)) + 
    geom_text(data = n_per_week_labels012, aes(label = label)) + 
    
    geom_vline(xintercept = ymd(c("2024-09-12", "2025-02-13", "2025-06-12")), 
               linetype = "dashed", color = "gray75") + 
    geom_text(data = n_per_week_batch_labels, aes(label = label), 
              color = "gray75", hjust = 0, nudge_x = 2) + 
    
    geom_point() + 
    geom_path() + 
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, c(0.01, 0.02)),
                       breaks = seq(0, 1, 0.1),
                       # minor_breaks = seq(0, 1, 0.02),
                       minor_breaks = NULL,
                       labels = scales::percent) +
    scale_color_manual(breaks = c("0", "1", "2+"),
                       # Color brewer Set 3
                       values = c("gray66", "#fb8072",  "#80b1d3")) +
    labs(x = NULL,
         y = "percentage of wards",
         title = "Number of new hymns (from all batches) sung by week",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), 
                          "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

The gray line here is the same as what was shown in the previous plot since it shows the proportion of wards singing zero new hymns per week. But when we collapse all batches together, we see a pretty good trend where more and more wards are singing one new hymn per week. It ebbed and flowed, and especially in the May--July era there were dips as we hit various holidays. But, I wonder if we'll see this continue going forward. 

What I notice too is that there seem to be increases in the number of wards singing new hymns a few weeks after the the release of a new hymn. This is most apparent after the second batch came out. During that time, we also see slightly more wards singing two or more new hymns each week. A few weeks after the third batch came out, there is another rise. And even maybe a month or so after the fourtch batch came out there was a slight increase, to the point where we hit that threshold of more wards singing new hymns than wards not singing new hymns. 

I want to zoom in briefly on the number singing two or more. The following plot shows the line labeled "2+" in the above plot, but broken down further into 2, 3, and 4 or more. Notice that this plot only goes up to about 6% in the *y*-axis, so we're really zoomed into the bottom portion of the previous plot.

```{r}
n_new_by_week234 <- meetings_since_release1 |> 
    mutate(is_new = hymn_num > 1000) |> 
    summarize(n_new = sum(is_new), .by = c(date, meeting_id)) |> 
    
    mutate(n_new = if_else(n_new >= 4, "4+", as.character(n_new))) |> 
    count(date, n_new) |>
    mutate(prop = n/sum(n), .by = date) |> 
    filter(n_new %in% c("2", "3", "4+")) |>
    print()
n_per_week_labels234 <- tibble(date = ymd(c("2025-01-10", "2024-12-15", "2024-11-20")),
                               prop = c(0.06, 0.03, 0.02),
                               n_new = c("2", "3", "4+"),
                               label = c("two new hymns", "three new hymn", "four or more new hymns"))
n_per_week_batch_labels <- n_per_week_batch_labels |> 
    mutate(prop = 0.05)
```



```{r, include = TRUE, fig.height = 4.5, fig.width = 14}
#| column: page
ggplot(n_new_by_week234, aes(date, prop, color = n_new)) + 
    
    geom_vline(xintercept = ymd(c("2024-09-12", "2025-02-13", "2025-06-12")),
               linetype = "dashed", color = "gray75") +
    geom_text(data = n_per_week_batch_labels, aes(label = label),
              color = "gray75", hjust = 0, nudge_x = 2) +
    
    geom_point() +
    geom_path() +
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, c(0.001, 0.005)),
                       breaks = seq(0, 1, 0.01),
                       # minor_breaks = seq(0, 1, 0.005),
                       minor_breaks = NULL,
                       labels = scales::percent) +
    scale_color_manual(breaks = c("2", "3", "4+"),
                       # Color brewer Set 3
                       values = c("#80b1d3", "#b3de69",  "#fdb462")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Number of new hymns (from all batches) sung by week",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), 
                          "sacrament meetings from", comma(n_wards_since_release1), "wards")) + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "bottom")
```

The biggest thing that I notice here is that the number of wards singing two new hymns really peaked in the three months after the second batch came out. I think people were pretty excited about another batch, and rather than singing the second batch *instead* of the first batch, they were adding it to their programs. At its peak, 5% of wards sang two new hymns. There was another brief period in May where we saw the same thing, but I'm not sure why that is. 

When we look at wards singing three or four new hymns, it's never been particularly common. Other than probably a fluke in early November, no more than 1% of wards have sung three new hymns. And I've never seen more than two wards in a week sing four new ones, and never since the fourth batch came out. 

## What hymns are most popular?

The following table shows the hymns from the fourth batch in order of how popular they are. On the right, I have an approximation of how many wards have sung that hymn during this period. I've seen all but <hymn>Still, Still, Still (#1207)</hymn> which makes sense since it's a Christmas hymn. 

```{r, include = TRUE}
meetings_since_release4 |>

    # Because people are repeating 1001 now, I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |>
    filter(nth_time == 1) |>

    cleveland(return = "table",
              stat_to_plot = sum_prop_wards,
              hymns_to_plot = c(1032:1041, 1207)) |>
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |>
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |>
    filter(hymn_num %in% c(1032:1041, 1207)) |>
    mutate(sum_prop_wards = scales::percent(round(sum_prop_wards, 3))) |>

    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           sum_prop_wards = if_else(is.na(sum_prop_wards), "0%", sum_prop_wards)) |>
    select(Hymn = label, `percent of wards` = sum_prop_wards) |>

    gt() |>
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Since June 15, 2025")
```

It's perhaps not too surprising that the most popular one is <hymn>O Lord, Who Gave Thy Life for Me (#1041)</hymn> since it's a sacrament hymn. For a while, <hymn>I'm a Pioneer Too (#1034)</hymn> was the #1 hymn in this batch, but keep in mind that it came out just before [Pioneer Day](/posts/pioneer_day) in Utah so it was briefly pretty popular. After a few weeks though, it sunk a little lower because I don't think too many wards will be singing it "off-season." I wonder if <hymn>Oh, How Great Is Our Joy (#1033)</hymn> is next most popular from this batch because the tune is (from what I understand) a well-known hymn in Spanish. I personally really like <hymn>Look unto Christ (#1032)</hymn> and am happy to see it near the top of this list because it means I'm likely to hear it every so often in the future. I'm also very pleased to see <hymn>I'm Gonna Live So God Can Use Me (#1037)</hymn> being sung as well. It's great to incorporate music from different backgrounds into our meetings.


## What hymns are most popular each week?

It's been long enough that we can get an idea of which hymns were popular when. The following plot shows the percentage of wards that sang each hymn *of the wards that sang any fourth batch hymns*. 

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 7}
#| column: page
meetings_since_release4 |> 
    filter(hymn_num %in% c(1032:1041, 1207)) |>
    count(date, hymn_num, name_num) |> 
    
    # pivot and stuff to get zeros
    pivot_wider(names_from = date, values_from = n, values_fill = 0) |> 
    pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
    mutate(date = ymd(date)) |> 
    
    # put them in order
    arrange(hymn_num) |> 
    mutate(name_num = fct_inorder(name_num)) |> 
    
    # Put the dates in order
    arrange(date) |> 
    mutate(date = date) |> 
    
    # proportions per week
    mutate(prop = n/sum(n), .by = date) |> 
    
    ggplot(aes(date, prop, group = name_num)) + 
    geom_path() + 
    geom_point() + 
    scale_x_date(date_breaks = "3 weeks",
                 date_minor_breaks = "1 week",
                 date_labels = "%b %d",
                 expand = expansion(0.1, 0)) +
    scale_y_continuous(labels = scales::percent) + 
    facet_wrap(~name_num, ncol = 4) + 
    labs(title = "Proportion of wards singing new hymns by week",
         subtitle = "Of the sacrament meetings with new hymns",
         x = NULL,
         y = "proportion of wards") + 
    theme_bw()
```

First, it's important to remember that hardly anyone sang from this batch during the first few weeks. So, the huge spike in <hymn>Because (#1039)</hymn> and <hymn>I'm Gonnna Live So God Can Use Me (#1037)</hymn> are mostly due to data sparsity issues. However, one true pattern was the big spike that <hymn>I'm a Pioneer Too (#1034)</hymn> experienced at the end of July, which is the Sunday before Pioneer Day here in Utah. On July 20th, half of wards that sang from the fourth batch sang that hymn that day. Other than those, I don't really notice any other patterns for any of the hymns. 

## When during meetings are these hymns sung?

We can also see when during sacrament meetings these hymns have been sung. I include this table for every batch, but it's tricky to get much meaningful information from it because every hymn has its own story. For reasons that are not always clear, some hymns just tend to be opening hymns, others tend to be closing, and others show some other perhaps more complicated pattern. 

```{r, include = TRUE}
meetings_since_release4 |> 
    filter(hymn_num %in% c(1032:1041, 1207),
           !is.na(type)) |>
    count(name_num, hymn_num, type) |> 
    mutate(n = round(n/sum(n), 2)*100, 
           .by = name_num) |> 
    pivot_wider(names_from = type, values_from = n, values_fill = 0) |>
    arrange(hymn_num) |> 
    select(-hymn_num) |> 
    select(name_num, 
           Opening, Sacrament, Intermediate, Closing) |> 
    rename(`New Hymn` = name_num) |> 
    gt() |> 
    gt_plt_bar_pct(column = c(Opening), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Sacrament), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Intermediate), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Closing), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Since June 15, 2025") |> 
    tab_source_note("Note: Each row adds up to 100%.")
```


It makes sense that hymn>O Lord, Who Gave Thy Life for Me (#1041)</hymn> is most commonly sung as a sacrament hymn because it is tagged as such. But, <hymn>As I Keep the Sabbath Day (#1035)</hymn> and <hymn>Look unto Christ (#1032)</hymn> were both occasionally sung as sacrament hymns. 

The next clearest trend is that <hymn>Read the Book of Mormon and Pray (#1036)</hymn> seems to be most common as an intermediate hymn. The one most commonly sung as an opening hymn from this batch was <hymn>As I Keep the Sabbath Day (#1035)</hymn> and the one most commonly sung as a closing hymn was <hymn>The Lord's My Shepherd (#1038)</hymn>. 


## Conclusion

So, that's it for the fourth batch of new hymns. The general pattern I see is that this batch has been less popular. It is true that it's smaller than the third batch, but it is still bigger than the second batch, so we can't blame numbers alone. We've also had some holidays here and there, like Father's Day, 4th of July, and Pioneer Day, so those likely take precedent. It may also just be that the novelty of new hymns has worn off, although it's important to note that almost half of wards are singing at least one new hymn each week. So maybe it's that we're still getting used to the hymns in the first batches before moving on to this latest one. 
