---
title: "The Fourth Batch of New Hymns!"
author: "Joey Stanley"
date: "2025-06-15"
date-modified: last-modified
categories: 
  - general
  - frequency
  - new hymns
reference-location: margin
---

```{r, include = FALSE}
knitr::opts_chunk$set(include = FALSE,
                      echo = FALSE,
                      fig.width = 8)
rmarkdown::render("../../_scripts/analysis_functions.Rmd")

# Apply a date stamp
sf <- stamp("June 2", quiet = TRUE)

# get last Sunday
last_sunday <- ymd(floor_date(today(), "week"))
library(gtExtras)
```



```{r}
meetings_since_release1 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-06-02"),
           !date %in% ymd("2024-10-06",
                          "2025-04-06"))
meetings_since_release2 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2024-09-15"),
           !date %in% ymd("2024-10-06",
                        "2025-04-06"))
meetings_since_release3 <- freqs |> 
    filter(date <= ymd("2025-06-08"),
           date >= ymd("2025-02-16"),
           !date %in% ymd("2025-04-06"))
meetings_since_release4 <- freqs |> 
    filter(date <= last_sunday,
           date >= ymd("2025-06-15"))

n_meetings_since_release1 <- meetings_since_release1 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release2 <- meetings_since_release2 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release3 <- meetings_since_release3 |> 
    distinct(meeting_id) |> 
    nrow()
n_meetings_since_release4 <- meetings_since_release4 |> 
    distinct(meeting_id) |> 
    nrow()

n_wards_since_release1 <- meetings_since_release1 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release2 <- meetings_since_release2 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release3 <- meetings_since_release3 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()
n_wards_since_release4 <- meetings_since_release4 |> 
    filter(date <= ymd(last_sunday)) |> 
    pull(ward) |> 
    unique() |> 
    length()

n_weeks_since_release1 <- meetings_since_release1 |> 
    count(date) |> 
    nrow()
n_weeks_since_release2 <- meetings_since_release2 |> 
    count(date) |> 
    nrow()
n_weeks_since_release3 <- meetings_since_release3 |> 
    count(date) |> 
    nrow()
n_weeks_since_release4 <- meetings_since_release4 |> 
    count(date) |> 
    nrow()
```


On June 12th, 2025, the church [released](https://newsroom.churchofjesuschrist.org/article/eleven-new-hymns-now-available-in-hymns-for-home-and-church) the fourth batch of new hymns. In previous posts, I have covered in detail the [first](/posts/new_hymns_batch1), [second](/posts/new_hymns_batch2), and [third](/posts/new_hymns_batch3) batches of hymns and how they were incorporated into sacrament meetings. This page covers period since when the fourth batch came out and will update weekly as more data comes in. <!--Currently I have data from `r scales::comma(n_meetings_since_release4)` sacrament meetings from `r scales::comma(n_wards_since_release4)` wards since June 15, 2025.-->


I'll tell you right now, the results are pretty boring. I have data from 576 wards and just one of them sang from the fourth batch. The hymn they sang was <hymn>Because (#1039)</hymn>. Honestly, I fully expected this one to be among the least common congregational hymns from this batch, so this is a surprise!


<!--

So, I'll continue with the stats anyway, here's how much data I have for each week so far, just so you have an idea of what I'm working with.

```{r, include = TRUE}
meetings_since_release4 |> 
    distinct(date, meeting_id) |> 
    count(date) |> 
    mutate(date = sf(date)) |> 
    rename(wards = n) |> 
    gt() |> 
    tab_header(title = "Number of wards I have data from, by week",
               subtitle = "Since June 15th, 2025")
```



## How many wards sang new hymns each week?

The following plot show what percentage of wards sang from each of the new batches of hymns, per week, since June 2024 when the new hymns started getting rolled out. I've included data from all three batches just so you can compare them to each other.

```{r, fig.height = 4, fig.width = 8}
count_new_hymns <- function(.category_name, .hymns) {
    meetings_since_release1 |> 
        mutate(is_new = hymn_num %in% .hymns) |>
        summarize(n_new = sum(is_new), .by = c(date, ward)) |>
        count(date, n_new) |>
        arrange(date) |>
        mutate(prop_new = n/sum(n), .by = date) |>
        mutate(n_new = as.factor(n_new)) |>
        mutate(category = .category_name) |>
        select(-n) |>
        pivot_wider(names_from = n_new, values_from = prop_new)
}
data_to_plot <- meetings_since_release1 |> 
    summarize(batch1 = sum(hymn_num %in% c(1001:1009, 1201:1204)) > 0,
              batch2 = sum(hymn_num %in% c(1010:1018)) > 0,
              batch3 = sum(hymn_num %in% c(1019:1031, 1205:1206)) > 0,
              batch4 = sum(hymn_num %in% c(1032:1041, 1207)) > 0,
              .by = c(meeting_id, date)) |>
    mutate(none = !batch1 & !batch2 & !batch3 & !batch4) |>
    pivot_longer(cols = c(batch1, batch2, batch3, batch4, none), names_to = "batch", values_to = "present") |> 
    count(date, batch, present) |> 
    pivot_wider(names_from = present, values_from = n, values_fill = 0) |> 
    mutate(prop_sung = `TRUE`/(`TRUE` + `FALSE`)) |> 
    
    filter(!(batch == "batch4" & date < ymd("2025-06-15"))) |>
    # don't need batch 3 before June 15
    filter(!(batch == "batch3" & date < ymd("2025-02-13"))) |>
    # don't need batch 2 before Sep 15
    filter(!(batch == "batch2" & date < ymd("2024-09-15"))) |>
    # don't need batch 1 before Sep 15 either
    filter(!(batch == "batch1" & date < ymd("2024-06-02"))) |>
    print()
```



```{r, fig.height = 4, fig.width = 8}
# Get date breaks for this post.
date_minor_breaks <- tibble(weeks_since = 0:52,
                            date = last_sunday) |> 
    mutate(date = last_sunday - weeks_since*7) |> 
    filter(date >= "2024-06-02")
date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 3 == 0)
batch4_date_breaks <- date_minor_breaks |> 
    filter(weeks_since %% 1 == 0)
```


```{r}
labels <- tibble(
    batch = c("none", "batch1", "batch2", "batch3", "batch4"),
    prop  = c(0.73, 0.26, 0.08, 0.04, 0.05),
    date  = ymd(c("2025-04-01", "2025-02-14", "2025-01-25", "2025-03-15", "2025-06-15")),
    label = c("no new hymns", "batch 1", "batch 2", "batch 3", "batch4"))

# Annotations
christmas_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2024-12-01",
           date <= "2025-01-01") |> 
    mutate(label = "Christmas")
easter_data_to_highlight <- data_to_plot |> 
    filter(batch %in% c("batch1", "batch2"),
           date >= "2025-04-13",
           date <= "2025-04-27") |> 
    mutate(label = "Easter")
```


```{r, include = TRUE, fig.height = 5.5, fig.width = 14}
#| column: page
ggplot(data_to_plot, aes(date, prop_sung, group = batch, color = batch)) +
    
    # Christmas annotation
    ggforce::geom_mark_ellipse(data = christmas_data_to_highlight, group = 1, color = "gray50") + 
    annotate(geom = "text", x = ymd("2024-11-03"), y = 0.35, label = "Christmas", hjust = 1, color = "gray50") + 
    annotate(geom = "curve", color = "gray50",
             x = ymd("2024-11-05"), xend = ymd("2024-11-25"), y = 0.35, yend = 0.30,
             curvature = -0.25, arrow = arrow(length = unit(2, "mm"))) + 
    
    # Easter
    ggforce::geom_mark_ellipse(data = easter_data_to_highlight, group = 1, color = "gray50") +
    annotate(geom = "segment", color = "gray50",
             x = ymd("2025-04-20"), xend = ymd("2025-04-20"), y = 0.35, yend = 0.275, 
             arrow = arrow(length = unit(2, "mm"))) +
    annotate(geom = "text", x = ymd("2025-04-20"), y = 0.37, label = "Easter", hjust = 0.5, color = "gray50") +
    
    # Mother's Day
    annotate(geom = "segment", 
             x = ymd("2025-05-11"), xend = ymd("2025-05-11"), y = 0.28, yend = 0.18, 
             arrow = arrow(length = unit(2, "mm")), color = "gray50") +
    annotate(geom = "text", x = ymd("2025-05-11"), y = 0.29, label = "Mother's\nDay", 
             vjust = 0, hjust = 0.5, lineheight = 0.75, color = "gray50") +
    
    
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(data = labels, aes(y = prop, label = label), hjust = 0.5, vjust = 0.5, size = 6) +
    
    
    scale_x_date(breaks = date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.02, 0)) +
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4"),
                       # Color brewer Set1
                       values = c("gray66", "#e41a1c",  "#377eb8",  "#4daf4a", "#984ea3")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since June 2",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards"),
         fill = "date") + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

```{r}
as_percent_joey <- function(x){
    paste0(round(x * 100, 1), "%")
}
```

Since it's only been one week, there's not a lot to say, almost nobody sang from the fourth batch. In fairness, it was Father's Day in the United States, so people likely had other plans. 

Here's another version of the same plot but instead of the actual dates, it's in terms of how many weeks since the batches came out. We can use this plot to spot any parallel trends across the batches, especially in the first few weeks of their release.

```{r, include = TRUE, fig.height = 5.5, fig.width = 14}
#| column: page
data_to_plot |> 
    mutate(weeks_since_release = as.numeric(date - min(date))/7, .by = batch) |> 
    filter(!batch %in% "none") |> 
    ggplot(aes(weeks_since_release, prop_sung, group = batch, color = batch)) +
    geom_path(linewidth = 1) +
    geom_point(size = 3) +
    scale_x_continuous(breaks = seq(2, 100, 2),
                       expand = expansion(0, c(0.2, 2))) + 
    scale_y_continuous(expand = expansion(0, 0.02),
                       breaks = seq(0, 1, 0.1),
                       minor_breaks = seq(0, 1, 0.02),
                       labels = scales::percent) +
    scale_color_manual(breaks = c("none", "batch1", "batch2", "batch3", "batch4"),
                       # Color brewer Set 1
                       values = c("gray66", "#e41a1c",  "#377eb8",  "#4daf4a", "#984ea3")) +
    labs(x = NULL,
         y = "percentage of wards",
         color = "number of new hymns",
         title = "Percentage of wards singing new hymns since their release date",
         subtitle = paste("Based on data from", comma(n_meetings_since_release1), "sacrament meetings from", comma(n_wards_since_release1), "wards"),
         fill = "date") + 
    theme_minimal(base_size = 16) +
    theme(legend.position = "none")
```

Here we can see that each batch is a little less popular than the previous one, both overall and week by week. Nearly 20% of wards sang from the first batch the week it came out. About 8% did for the second batch. About 1% for the third batch. And of the 576 wards I have data from on June 15th, just one of them sang from the fourth batch. 


```{r}
data_to_plot |> 
    filter(batch == "batch4")
```

<!--
## What hymns are most popular?

The following table shows the hymns from the third batch in order of how popular they are. On the right, I have an approximation of how many wards have sung that hymn during this period. 

```{r, include = TRUE, eval = FALSE}
meetings_since_release4 |> 
    
    # Because people are repeating 1001 now, I can't take it as the raw frequency. I'll have to remove duplicates.
    mutate(nth_time = row_number(), .by = c(ward, hymn_num)) |> 
    filter(nth_time == 1) |> 
    
    cleveland(return = "table", 
              stat_to_plot = sum_prop_wards,
              hymns_to_plot = c(1032:1041, 1207)) |> 
    mutate(hymn_num = as.numeric(as.character(hymn_num))) |> 
    full_join(name_num_lookup, by = c("hymn_name", "hymn_num")) |> 
    filter(hymn_num %in% c(1032:1041, 1207)) |>
    mutate(sum_prop_wards = scales::percent(round(sum_prop_wards, 3))) |> 
    
    # add label for missing hymns
    mutate(label = paste0(hymn_name, " (", hymn_num, ")"),
           sum_prop_wards = if_else(is.na(sum_prop_wards), "0%", sum_prop_wards)) |> 
    select(Hymn = label, `percent of wards` = sum_prop_wards) |> 
    
    gt() |> 
    tab_header(title = "Approximate percentage of wards new hymns have been sung in",
               subtitle = "Since June 15, 2025")
```

The only hymn that was sung was <hymn>Because (#1039)</hymn>. It's not showing up on this plot because of rounding issues. Honestly, I fully expected this one to be among the least common congregational hymns from this batch, so this is a surprise! 


## What hymns are most popular each week?

It's been long enough that we can get an idea of which hymns were popular when. The following plot shows the percentage of wards that sang each hymn *of the wards that sang any third batch hymns*. 

```{r, include = TRUE, warning = FALSE, message = FALSE, fig.height = 7, fig.width = 13, eval = FALSE}
#| column: page
meetings_since_release3 |> 
    filter(hymn_num %in% c(1019:1031, 1205:1206)) |>
    count(date, hymn_num, name_num) |> 
    
    # pivot and stuff to get zeros
    pivot_wider(names_from = date, values_from = n, values_fill = 0) |> 
    pivot_longer(cols = -c(hymn_num, name_num), names_to = "date", values_to = "n") |>
    mutate(date = ymd(date)) |> 
    
    # put them in order
    arrange(hymn_num) |> 
    mutate(name_num = fct_inorder(name_num)) |> 
    
    # Put the dates in order
    arrange(date) |> 
    mutate(date = date) |> 
    
    # proportions per week
    mutate(prop = n/sum(n), .by = date) |> 
    
    ggplot(aes(date, prop, group = name_num)) + 
    geom_path() + 
    geom_point() + 
    scale_x_date(breaks = batch3_date_breaks$date,
                 minor_breaks = date_minor_breaks$date,
                 date_labels = "%b %d",
                 expand = expansion(0.1, 0)) +
    scale_y_continuous(labels = scales::percent) + 
    facet_wrap(~name_num, ncol = 4) + 
    labs(title = "Proportion of wards singing new hymns by week",
         subtitle = "Of the sacrament meetings with new hymns",
         x = NULL,
         y = "proportion of wards") + 
    theme_bw()
```

The most obvious trend though is <hymn>Let Easter Anthems Ring (#1205)</hymn> and <hymn>Were You There? (#1206)</hymn> being sung around [Easter](/posts/easter2025). Other than that, there isn't much of a pattern for these hymns. Some of the more popular ones jumped around in popularity a little bit from week to week, but it's not like we're seeing an obvious trend of one hymn being sung more often soon after the release of the batch---as if wards were anxious to sing it---like what we saw in the [first batch](/posts/new_hymns_batch1)

## When during meetings are these hymns sung?

We can now see when during sacrament meetings these hymns have been sung. 
```{r, include = TRUE, eval = FALSE}
meetings_since_release3 |> 
    filter(hymn_num %in% c(1019:1031, 1205:1206),
           !is.na(type)) |>
    count(name_num, hymn_num, type) |> 
    mutate(n = round(n/sum(n), 2)*100, 
           .by = name_num) |> 
    pivot_wider(names_from = type, values_from = n, values_fill = 0) |>
    arrange(hymn_num) |> 
    select(-hymn_num) |> 
    select(name_num, Opening, Sacrament, Intermediate, Closing) |> 
    rename(`New Hymn` = name_num) |> 
    gt() |> 
    gt_plt_bar_pct(column = c(Opening), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |> 
    gt_plt_bar_pct(column = c(Sacrament), scaled = TRUE, labels = TRUE,
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Intermediate), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    gt_plt_bar_pct(column = c(Closing), scaled = TRUE, labels = TRUE, 
                   fill = "#4F2B47", height = 20, font_size = "14px") |>
    tab_header(title = "When were new hymns sung in sacrament meeting?",
               subtitle = "Since February 15") |> 
    tab_source_note("Note: Each row adds up to 100%.")
```

Again, like any plot like this, each hymn has its own story it seems. Some are simply more popular as opening hymns, or closing hymns, or intermediate hymns. We didn't really get any sacrament hymns in this batch, but that didn't stop people from singing some from this batch very occasionally. It seems like most of these hymns were pretty evenly spread out between being sung as opening, intermediate, and closing hymns. 

-->

## Conclusion

There's really not much more to say right now! We'll see if more people sing from this batch next week!
